---
title: "Streamlined Pacific Cod Analyses & Figures"
author: "Laura Vary"
date: "1/23/2022"
header-includes: 
  - \usepackage{placeins}
output: 
  pdf_document: 
    toc: yes
---

```{r setup, echo=FALSE,message=FALSE,warning=FALSE,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE,message=FALSE,warning=FALSE,include=FALSE}
setwd("C:/Users/varyl/Desktop/GitHub/BeringSeaFishes_GAM_Analyses")
library(knitr)
library(raster)
library(marmap)
library(maps)
library(mgcv)
library(colorRamps) 
library(RColorBrewer)
library(dplyr)
library(fields)
library(mapdata)
library(ggplot2)
library(ggmap)
library(ggpubr)
library(scales)
source('./Organized Thesis Analyses (Code)/distance.function.R')
source("./Organized Thesis Analyses (Code)/vis.gam_COLORS.R")
source("./Organized Thesis Analyses (Code)/euclidean.distance.R")
```

# Pacific Cod (Pac cod): 

## Loading Data: 

Pacific cod: larval data only are included for this species. Pac cod spawn in March and April under the sea ice, live roughly 13 years, and transform to juveniles at standard lengths between 25 and 35 mm. 

```{r,echo=FALSE}
pclarv.ctd<-read.csv('./Ichthyo Data/Cleaned_Cut_PcLarv_wCTD.csv',header=TRUE,check.names=TRUE)
pclarv.ctd<-subset(pclarv.ctd,lat<60)
pclarv.ctd<-subset(pclarv.ctd,doy>99&doy<160)

str_name<-"./Environmental Data/expanded_BS_bathy.tif"
bathy<-raster(str_name)

bathybath<-as.bathy(bathy)
```

These data have been trimmed. The larval data are constrained to depths between 40 and 250 meters and to latitudes south of 62 degrees north. Larvae are linked to CTD-derived, $in$ $situ$ temperature and salinity measurements and constrained to days of year between 99 and 160.

## Descriptive Information: 

```{r,echo=FALSE}
lat_range<-c(paste(round(min(pclarv.ctd$lat),digits=1),round(max(pclarv.ctd$lat),digits=1),sep="-"))
lon_range<-c(paste(round(min(pclarv.ctd$lon),digits=1),round(max(pclarv.ctd$lon),digits=1),sep=" to "))
doy_range<-c(paste(min(pclarv.ctd$doy),max(pclarv.ctd$doy),sep="-"))
depth_range<-c(paste(round(min(pclarv.ctd$bottom_depth),digits=1),
                     round(max(pclarv.ctd$bottom_depth),digits=1),sep="-"))

df<-data.frame(lat_range,lon_range,doy_range,depth_range)

kable(df,col.names=c("Lat Range","Lon Range","Day of Year Range",
                     "Bottom Depth Range"),
      caption="Descriptive Metrics for Pac Cod Larval Data",align='c')
```

The following two plots show *the day of year distribution for positive Pac cod larval catch* (left) and *the year distribution for positive Pac cod larval catch* (right). 

```{r, fig.width=7.5,fig.height=3.5,fig.fullwidth=TRUE,fig.cap="Pac Cod Larvae", echo=FALSE}
par(mfrow=c(1,2))
hist(pclarv.ctd$doy[pclarv.ctd$Cper10m2>0],xlab="Day of Year",ylab="Positive Catch Frequencies",main="")
hist(pclarv.ctd$year[pclarv.ctd$Cper10m2>0],xlab="Year",ylab="",main="")
```
\newpage

The following plots show Pacific cod larval catch distributions (Catch per unit effort, or per 10m$^2$) across five year increments from 1997 to 2016 
```{r,echo=FALSE}
pcylarv<-pclarv.ctd%>%mutate(yearbin=case_when(year<=2002~'1997-2002',
                                               year<=2007&year>2002~'2003-2007',
                                               year<=2012&year>2007~'2008-2012',
                                               year<=2017&year>2012~'2013-2016'))

pcylpos<-pcylarv[pcylarv$Cper10m2>0,]
pcylpos$logcpue<-log(pcylpos$Cper10m2+1)
pcylz<-pcylarv[pcylarv$Cper10m2==0,]
```

```{r,fig.height=10,fig.width=8,fig.align='center',fig.fullwidth=T,echo=FALSE,warning=FALSE}
autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=pcylz,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=pcylpos,aes(x=lon,y=lat,size=logcpue),
                               fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous("Log (CPUE+1)",range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  scale_shape_manual(values=('No Catch'=8))+
  facet_wrap(~yearbin)+
  labs(title="Pacific Cod Sole Larvae")+ylab("Latitude")+xlab("Longitude")
```


## Larval Generalized Additive Models: 

Now we'll move into the GAMs. The following code *is only necessary if the data were re-trimmed and new GAMs need to be run*. In this case, modify markdown document such that "{eval = TRUE}". The other model figures are marked as "eval = FALSE" if they, as of the last model run, do not produce the best model results. **Make sure to save the new models as RDS objects.** 

Pac cod larvae were best explained by the bivariate salinity-temperature model, in which the spatial and temporal distribution of larvae were modelled in association with a smooth containing *in situ* salinity-temperature data. 

We begin with the base larval model: 
```{r,eval=F,echo=FALSE}
lv.base<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
               s(bottom_depth,k=5),
             data=pclarv.ctd,family=tw(link='log'),method='REML')

saveRDS(lv.base,"./GAM Models/pc_larvae_base.rds")
```

```{r}
lv.base<-readRDS("./GAM Models/pc_larvae_base.rds")
summary(lv.base)
AIC(lv.base)
```

```{r,eval=F,echo=FALSE}
lv.add.sal<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                  s(bottom_depth,k=5)+
                  s(salinity),data=pclarv.ctd,family=tw(link='log'),
                method='REML')

saveRDS(lv.add.sal,file="./GAM Models/pc_larvae_addsal.rds")
```

```{r,include=FALSE}
lv.add.sal<-readRDS("./GAM Models/pc_larvae_addsal.rds")
summary(lv.add.sal)
AIC(lv.add.sal)
```

```{r,eval=F,echo=FALSE}
lv.add.temp<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                   s(bottom_depth,k=5)+
                   s(temperature),data=pclarv.ctd,family=tw(link='log'),
                 method='REML')

saveRDS(lv.add.temp,file="./GAM Models/pc_larvae_addtemp.rds")
```

```{r,include=FALSE,echo=FALSE}
lv.add.temp<-readRDS("./GAM Models/pc_larvae_addtemp.rds")
summary(lv.add.temp)
AIC(lv.add.temp)
```

Then additive temperature and salinity, in individual additive terms. This is the second-best performing model. 
```{r,eval=F,echo=FALSE}
lv.temp.sal<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                   s(bottom_depth,k=5)+
                   s(temperature)+s(salinity),data=pclarv.ctd,
                 family=tw(link='log'),method='REML')

saveRDS(lv.temp.sal,file="./GAM Models/pc_larvae_addtempsal.rds")
```

```{r}
lv.temp.sal<-readRDS("./GAM Models/pc_larvae_addtempsal.rds")
summary(lv.temp.sal)
AIC(lv.temp.sal)
```

And finally, the best performing model: the bivariate salinity-temperature additive term: 
```{r,eval=F,echo=FALSE}
lv.2d<-gam((Cper10m2+1)~factor(year)+s(lon,lat)+s(doy,k=7)+s(bottom_depth)+
             te(salinity,temperature),data=pclarv.ctd,family=tw(link='log'),
           method='REML')

saveRDS(lv.2d,file="./GAM Models/pc_larvae_2d.rds")
```

```{r}
lv.2d<-readRDS("./GAM Models/pc_larvae_2d.rds")
summary(lv.2d)
AIC(lv.2d)
```

To confirm that this is indeed the best model, we can compare AIC values across all five tested models. 
```{r, echo=FALSE}
aic.base<-AIC(lv.base)
aic.sal<-AIC(lv.add.sal)
aic.temp<-AIC(lv.add.temp)
aic.saltemp<-AIC(lv.temp.sal)
aic.2d<-AIC(lv.2d)

aic.pclarv<-data.frame('model'=c('Base','Add Sal','Add Temp',
                                'Sal and Temp','2-d Sal-Temp'),
                      'AIC_value'=c(aic.base,aic.sal,aic.temp,
                                    aic.saltemp,aic.2d))
```



```{r,echo=FALSE}
windows(width=6,height=5);par(mai=c(1,1,0.5,0.9))
plot(c(1:5),aic.pclarv$AIC_value,main='PC',
     col=hcl.colors(5,"Zissou 1"),
     pch=19,cex=2,ylab='AIC Score',xlab='')
grid(nx=5,ny=14,col="lightgray")
text(c(1:5),aic.pclarv$AIC_value,labels=round(aic.pclarv$AIC_value),pos=c(4,1,3,2,2))
legend("bottomleft",legend=c('Base','Add. SSS','Add. SST',
                             'Add. SSS, SST','Bivariate SSS-SST'),
       col=hcl.colors(5,"Zissou 1"),
       lwd=3,lty=1,cex=0.7)
```



The following plot is the predicted Pac cod larval biogeography based on the best performing model, the bivariate salinity-temperature GAM. Observations (log transformed, n+1) are shown as well. 
```{r,echo=FALSE}
nlat=120
nlon=120
latd=seq(min(pclarv.ctd$lat,na.rm=TRUE),max(pclarv.ctd$lat,na.rm=TRUE),length.out=nlat)
lond=seq(min(pclarv.ctd$lon,na.rm=TRUE),max(pclarv.ctd$lon,na.rm=TRUE),length.out=nlon)

grid.extent<-expand.grid(lond,latd)
names(grid.extent)<-c('lon','lat')

grid.extent$dist<-NA
for(k in 1:nrow(grid.extent)){
  dist<-distance.function(grid.extent$lat[k],grid.extent$lon[k],
                          pclarv.ctd$lat,pclarv.ctd$lon)
  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-as.numeric(2005)
grid.extent$doy<-as.numeric(median(pclarv.ctd$doy,na.rm=TRUE))
grid.extent$bottom_depth<-NA
grid.extent$bottom_depth<-as.numeric(median(pclarv.ctd$bottom_depth,na.rm=TRUE))
grid.extent$temperature<-as.numeric(mean(pclarv.ctd$temperature))
grid.extent$salinity<-as.numeric(mean(pclarv.ctd$salinity))
grid.extent$pred<-predict(lv.2d,newdata=grid.extent)
grid.extent$pred[grid.extent$dist>30000]<-NA

symcol<-adjustcolor("grey",alpha=0.5)
```

```{r,fig.align='center',fig.fullwidth=TRUE,fig.width=6.5,fig.height=10,echo=FALSE}
par(mai=c(1,1,0.5,0.9),mfrow=c(2,1))
image.plot(lond,latd,t(matrix(grid.extent$pred,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=c(-180,-155),ylim=c(52,63),
           main='Predicted Larval Biogeography, 2D Model',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("Anomalies in (log(C/(10m"^2,')+1)')),legend.shrink=0.3)
symbols(pclarv.ctd$lon[pclarv.ctd$Cper10m2>0],pclarv.ctd$lat[pclarv.ctd$Cper10m2>0],
        circles=log(pclarv.ctd$Cper10m2+1)[pclarv.ctd$Cper10m2>0],inches=0.1,bg=symcol,fg='black',add=T)
points(pclarv.ctd$lon[pclarv.ctd$Cper10m2==0],pclarv.ctd$lat[pclarv.ctd$Cper10m2==0],
       pch="+",col="white")
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)

image.plot(lond,latd,t(matrix(grid.extent$pred,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=c(-180,-155),ylim=c(52,63),
           main='',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)
```

```{r,fig.align='center',fig.fullwidth=TRUE,fig.width=6.5,fig.height=5,echo=FALSE}
par(mai=c(1,1,0.6,0.9))
plot(lv.2d,select=2,shade=T,seWithMean=TRUE,shade.col="lightgrey",
     xlab="Day of Year",ylab="Day of Year Effect")
abline(h=0,col="mistyrose4",lty=2,lwd=1.3)
```

With this bivariate model, we can also calculate the predicted anomalous larval catch (more or less than expected) on a salinity-temperature plot. This figure shows that prediction, with observed larval catch (log(n=1)) overlaid. 
```{r,echo=FALSE}
ntemp<-100
nsal<-100
tempd<-seq(min(pclarv.ctd$temperature,na.rm=TRUE),max(pclarv.ctd$temperature,na.rm=TRUE),length.out=ntemp)
sald<-seq(min(pclarv.ctd$salinity,na.rm=T),max(pclarv.ctd$salinity,na.rm=T),length.out=nsal)

grid.extent<-expand.grid(sald,tempd)
names(grid.extent)<-c('salinity','temperature')

grid.extent$dist<-NA
for(k in 1:nrow(grid.extent)){
  dist<-euclidean.distance(grid.extent$salinity[k],grid.extent$temperature[k],
                               pclarv.ctd$salinity,pclarv.ctd$temperature)
  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-as.numeric(2005)
grid.extent$lon<-as.numeric(median(pclarv.ctd$lon))
grid.extent$lat<-as.numeric(median(pclarv.ctd$lat))
grid.extent$doy<-as.numeric(median(pclarv.ctd$doy,na.rm=TRUE))
grid.extent$bottom_depth<-NA
grid.extent$bottom_depth<-as.numeric(median(pclarv.ctd$bottom_depth,na.rm=TRUE))
grid.extent$pred<-predict(lv.2d,newdata=grid.extent)
grid.extent$pred[grid.extent$dist>mean(grid.extent$dist)]<-NA #threshold based on means
```

```{r,echo=FALSE,fig.align='center',fig.width=6.5,fig.height=10,fig.fullwidth=TRUE}
par(mai=c(1,1,0.5,0.9),mfrow=c(2,1))
image.plot(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
           col=hcl.colors(100,"Lajolla"),xlab='Salinity (psu)',
           ylab=expression(paste("Temperature ("^0, 'C)')),
           xlim=c(29,33.5),ylim=c(-1.8,14),
           main='',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("Anomalies in (log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
        col="white",levels=c(2,2.2),add=T)

image.plot(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
           col=hcl.colors(100,"Lajolla"),xlab='Salinity (psu)',
           ylab=expression(paste("Temperature ("^0, 'C)')),
           xlim=c(29,33.5),ylim=c(-1.8,14),
           main='Larval Biogeography By Temperature and Salinity',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("Anomalies in (log(C/(10m"^2,')+1)')),legend.shrink=0.3)
symbols(pclarv.ctd$salinity[pclarv.ctd$Cper10m2>0],
        pclarv.ctd$temperature[pclarv.ctd$Cper10m2>0],
        circles=log(pclarv.ctd$Cper10m2+1)[pclarv.ctd$Cper10m2>0],
        inches=0.1,bg=symcol,fg='black',add=T)
points(pclarv.ctd$salinity[pclarv.ctd$Cper10m2==0],pclarv.ctd$temperature[pclarv.ctd$Cper10m2==0],
       pch="+",col="white")
```

Now I'll calculate a specific range of temperature, salinity, and both temperature and salinity to evaluate breadth of environmental tolerances. 

For the univariate predictions, make a grid that holds either temperature or salinity constant depending on the variable of interest. The temperature and salinity values held constant were chosen based on the bivariate temperature,salinity plots; I attempted to capture a representative snapshot of variability in the univariate dimension. 

First, I'll determine the sum of larval log(CPUE+1) predictions and 60% of that sum, to know how many rows to cut from the larger extent. 

```{r,echo=FALSE}
sum<-sum(grid.extent$pred,na.rm=T)
sum

clip60<-0.6*sum
clip60 #determine 60% of summed CPUE predictions 

grid.extent1<-grid.extent
grid.extent1$pos<-seq.int(nrow(grid.extent1))
grid.extent1<-grid.extent1%>%arrange(pred)
grid.extent1$consum<-cumsum(grid.extent1$pred)
grid.extent1<-grid.extent1[grid.extent1$consum>clip60,]

grid.clip<-grid.extent1%>%arrange(grid.extent1$pos)

#write.csv(grid.clip,'Pc_larv_clip_60.csv')
```

So this percentage (percbi, 49.2%) is the percentage of the total grid extent within which 60% of the predicted log(CPUE+1) lies. I.e., 49.2% percent of the whole grid extent encompasses 60% of the predicted observations. 

```{r,echo=FALSE}
ntemp<-100
tempd<-seq(min(pclarv.ctd$temperature,na.rm=TRUE),max(pclarv.ctd$temperature,na.rm=TRUE),length.out=ntemp)

unit<-data.frame(matrix(ncol=1,nrow=100))
names(unit)<-'temperature'
unit$temperature<-as.numeric(tempd)

for(k in 1:nrow(unit)){
  unit$dist[k]<-unit$temperature[k]-pclarv.ctd$temperature[k]
}

unit$year<-as.numeric(2005)
unit$lon<-as.numeric(median(pclarv.ctd$lon))
unit$lat<-as.numeric(median(pclarv.ctd$lat))
unit$doy<-as.numeric(median(pclarv.ctd$doy,na.rm=TRUE))
unit$bottom_depth<-as.numeric(median(pclarv.ctd$bottom_depth,na.rm=TRUE))
unit$salinity<-31.7 #near the peak predictions 
unit$pred<-predict(lv.2d,newdata=unit)
unit$salinity<-31.3
unit$pred.1<-predict(lv.2d,newdata=unit)
unit$salinity<-30.8
unit$pred.2<-predict(lv.2d,newdata=unit)
unit$salinity<-32.1
unit$pred.3<-predict(lv.2d,newdata=unit)
unit$salinity<-32.5
unit$pred.4<-predict(lv.2d,newdata=unit)
unit$salinity<-30.5
unit$pred.5<-predict(lv.2d,newdata=unit)
unit$pred[unit$dist>mean(unit$dist)]<-NA
unit$pred.1[unit$dist>mean(unit$dist)]<-NA
unit$pred.2[unit$dist>mean(unit$dist)]<-NA
unit$pred.3[unit$dist>mean(unit$dist)]<-NA
unit$pred.4[unit$dist>mean(unit$dist)]<-NA
unit$pred.5[unit$dist>mean(unit$dist)]<-NA

nsal<-100
sald<-seq(min(pclarv.ctd$salinity,na.rm=T),max(pclarv.ctd$salinity,na.rm=T),length.out=nsal)
unis<-data.frame(matrix(ncol=1,nrow=100))
names(unis)<-'salinity'
unis$salinity<-as.numeric(sald)

for(k in 1:nrow(unis)){
  unis$dist[k]<-unis$salinity[k]-pclarv.ctd$salinity[k]
}

unis$year<-as.numeric(2005)
unis$lon<-as.numeric(median(pclarv.ctd$lon))
unis$lat<-as.numeric(median(pclarv.ctd$lat))
unis$doy<-as.numeric(median(pclarv.ctd$doy,na.rm=TRUE))
unis$bottom_depth<-as.numeric(median(pclarv.ctd$bottom_depth,na.rm=TRUE))
unis$temperature<-5.6 #centralized in peak predictions
unis$pred<-predict(lv.2d,newdata=unis)
unis$temperature<-5.8
unis$pred.1<-predict(lv.2d,newdata=unis)
unis$temperature<-6
unis$pred.2<-predict(lv.2d,newdata=unis)
unis$temperature<-6.3
unis$pred.3<-predict(lv.2d,newdata=unis)
unis$temperature<-5.3
unis$pred.4<-predict(lv.2d,newdata=unis)
unis$temperature<-5
unis$pred.5<-predict(lv.2d,newdata=unis)
unis$pred[unis$dist>mean(unis$dist)]<-NA
unis$pred.1[unis$dist>mean(unis$dist)]<-NA
unis$pred.2[unis$dist>mean(unis$dist)]<-NA
unis$pred.3[unis$dist>mean(unis$dist)]<-NA
unis$pred.4[unis$dist>mean(unis$dist)]<-NA
unis$pred.5[unis$dist>mean(unis$dist)]<-NA

unit<-unit[!is.na(unit$pred),]
unis<-unis[!is.na(unis$pred),]

clipt60<-0.6*(sum(unit$pred,na.rm=T))
clips60<-0.6*(sum(unis$pred,na.rm=T))

grid.t1<-unit
grid.t1$pos<-seq.int(nrow(grid.t1))
grid.t1<-grid.t1%>%arrange(pred)
grid.t1$consum<-cumsum(grid.t1$pred)
grid.t1<-grid.t1[grid.t1$consum>clipt60,]
grid.clip.t<-grid.t1%>%arrange(grid.t1$pos)

grid.s1<-unis
grid.s1$pos<-seq.int(nrow(grid.s1))
grid.s1<-grid.s1%>%arrange(pred)
grid.s1$consum<-cumsum(grid.s1$pred)
grid.s1<-grid.s1[grid.s1$consum>clips60,]
grid.clip.s<-grid.s1%>%arrange(grid.s1$pos)

```


```{r,echo=FALSE,fig.align='center',fig.width=8,fig.height=6,fig.fullwidth=TRUE,warning=FALSE}
temp<-ggplot(grid.clip.t,aes(x=temperature,y=pred))+geom_point(color='grey')+geom_smooth(method="loess")+
  theme_bw()+theme(axis.text.x=element_text(angle=45,hjust=1,size=7.5))+
  labs(x=expression(paste("Temperature ("^0,'C)')),y='Prediction')+
  geom_line(aes(x=temperature,y=pred.1),linetype="solid",color="grey48")+
  geom_line(aes(x=temperature,y=pred.2),linetype="twodash",color="grey48")+
  geom_line(aes(x=temperature,y=pred.3),linetype="longdash",color="grey48")+
  geom_line(aes(x=temperature,y=pred.4),linetype="dotted",color="grey48")+
  geom_line(aes(x=temperature,y=pred.5),linetype="dotdash",color="grey48")+
  scale_x_continuous(breaks=seq(-1.7,14,by=1),limits=c(-1.7,14))+
  scale_y_continuous(breaks=seq(1,7.3,by=1),limits=c(1,7.3))

sal<-ggplot(grid.clip.s,aes(x=salinity,y=pred))+geom_point(color='grey')+geom_smooth(method="loess")+
  theme_bw()+labs(x='Salinity (psu)',y='Prediction')+
  theme(axis.text.x=element_text(hjust=1,angle=45,size=7.5))+
  geom_line(aes(x=salinity,y=pred.1),linetype="solid",color="grey48")+
  geom_line(aes(x=salinity,y=pred.2),linetype="twodash",color="grey48")+
  geom_line(aes(x=salinity,y=pred.3),linetype="longdash",color="grey48")+
  geom_line(aes(x=salinity,y=pred.4),linetype="dotted",color="grey48")+
  geom_line(aes(x=salinity,y=pred.5),linetype="dotdash",color="grey48")+
  scale_x_continuous(breaks=round(seq(29,33.3,by=0.5),1),limits=c(29,33.3))+
  scale_y_continuous(breaks=seq(1,7.3,by=1),limits=c(1,7.3))

par(mai=c(1,1,0.5,0.9))
ggarrange(temp,sal)
```

For the bivariate analysis: 

```{r,echo=FALSE,fig.align='center',fig.width=6,fig.height=6,fig.fullwidth=TRUE}
par(mai=c(1,1,0.5,0.9))
bivar<-ggplot(grid.clip)+geom_tile(aes(x=salinity,y=temperature,fill=pred))+theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1,size=7.5))+
  geom_vline(xintercept=31.7,linetype="dashed",color="black")+
  geom_vline(xintercept=31.3,linetype="solid",color="grey48")+
  geom_vline(xintercept=30.8,linetype="twodash",color="grey48")+
  geom_vline(xintercept=32.1,linetype="longdash",color="grey48")+
  geom_vline(xintercept=32.5,linetype="dotted",color="grey48")+
  geom_vline(xintercept=30.5,linetype="dotdash",color="grey48")+
  geom_hline(yintercept=5.6,linetype="dashed",color="black")+
  geom_hline(yintercept=5.8,linetype="solid",color="grey48")+
  geom_hline(yintercept=6,linetype="twodash",color="grey48")+
  geom_hline(yintercept=6.3,linetype="longdash",color="grey48")+
  geom_hline(yintercept=5.3,linetype="dotted",color="grey48")+
  geom_hline(yintercept=5,linetype="dotdash",color="grey48")+
  labs(x='Salinity (psu)',y=expression(paste("Temperature ("^0,'C)')))+scale_fill_viridis(name='Prediction')+
  scale_x_continuous(breaks=seq(29,33.3,by=0.5),limits=c(29,33.3))+
  scale_y_continuous(breaks=seq(-1.7,14,by=1),limits=c(-1.7,14))
bivar
```

Determine area in T-S units over which 60% of observations are contained.

```{r,echo=FALSE}
width<-max(grid.extent$salinity)-min(grid.extent$salinity)
height<-max(grid.extent$temperature)-min(grid.extent$temperature)
area.all<-width*height

width.cl<-max(grid.clip$salinity)-min(grid.clip$salinity)
height.cl<-max(grid.clip$temperature)-min(grid.clip$temperature)
area.clip<-width.cl*height.cl

perc.area<-(area.clip/area.all)*100
perc.area
```

92% of the salinity-temperature area contains 60% of the total predicted log(CPUE+1). 

To again share the improvements of the best performing models from the base models, we can look at the AIC division produces. 
```{r,echo=FALSE}
bestdiv<-c(AIC(lv.2d)/AIC(lv.base))
second<-c(AIC(lv.2d)/AIC(lv.temp.sal))
stage<-"Larvae"

df<-data.frame(stage,bestdiv,second)
kable(df,col.names=c("","Best Divided By Base","Best Divided By Second Best"),
      caption="Model Power through AIC Comparisons, Pacific Cod",align='c')
```

Reduction in MSE (%):  
```{r,echo=FALSE}
var.ratio.lv<-(summary(lv.base)$scale-summary(lv.2d)$scale)/summary(lv.base)$scale
perc<-var.ratio.lv*100
print(perc)
```


## Appendix

The following plots investigate the diverging phenological peaks in larval density in order to evaluate whether different peaks correlate to different spatial areas in the BS. 

```{r,echo=FALSE}
pcd1<-pclarv.ctd[pclarv.ctd$Cper10m2>0&pclarv.ctd$doy>124&pclarv.ctd$doy<136,]
pcd1$logcpue<-log(pcd1$Cper10m2+1)
pcdz1<-pclarv.ctd[pclarv.ctd$Cper10m2==0&pclarv.ctd$doy>124&pclarv.ctd$doy<136,]

pcd2<-pclarv.ctd[pclarv.ctd$Cper10m2>0&pclarv.ctd$doy>144&pclarv.ctd$doy<161,]
pcd2$logcpue<-log(pcd2$Cper10m2+1)
pcdz2<-pclarv.ctd[pclarv.ctd$Cper10m2==0&pclarv.ctd$doy>144&pclarv.ctd$doy<161,]
```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
set1<-autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=pcdz1,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=pcd1,aes(x=lon,y=lat,size=logcpue),
                        fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous("Log (CPUE + 1)",range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  labs(title="Pac Cod: DOY 125-135",
       shape="Log (CPUE + 1)")+ylab("Latitude")+xlab("Longitude")

set2<-autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=pcdz2,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=pcd2,aes(x=lon,y=lat,size=logcpue),
                        fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous("Log (CPUE + 1)",range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  labs(title="DOY 145-160",
       shape="Log (CPUE + 1)")+ylab("Latitude")+xlab("Longitude")
```

```{r,fig.fullwidth=T,fig.align='center',fig.width=8,fig.height=10,echo=F}
ggarrange(set1,set2,common.legend=T,legend="bottom")
```

