---
title: "Streamlined Yellowfin Analyses & Figures"
author: "Laura Vary"
date: "1/25/2022"
header-includes: 
  - \usepackage{placeins}
output: 
  pdf_document: 
    toc: yes
---

```{r setup, echo=FALSE,message=FALSE,warning=FALSE,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE,message=FALSE,warning=FALSE,include=FALSE}
setwd("C:/Users/varyl/Desktop/GitHub/BeringSeaFishes_GAM_Analyses")
library(knitr)
library(raster)
library(marmap)
library(maps)
library(dplyr)
library(mgcv)
library(colorRamps) 
library(RColorBrewer)
library(fields)
library(mapdata)
library(ggplot2)
library(ggmap)
library(ggpubr)
library(scales)
source('./Organized Thesis Analyses (Code)/distance.function.R')
source("./Organized Thesis Analyses (Code)/vis.gam_COLORS.R")
source("./Organized Thesis Analyses (Code)/euclidean.distance.R")
```

# Yellowfin sole (YFS): 

## Loading Data: 

Yellowfin sole: larval data only are included for this species. Yellowfin sole spawn in April and May, live roughly 17 to 19 years, and transform to juveniles at standard lengths between 15 and 17 mm. 

```{r,echo=FALSE}
yflarv.ctd<-read.csv(file='./Ichthyo Data/Cleaned_Cut_YfLarv_wCTD.csv',
                     header=TRUE,check.names=TRUE)
yflarv.ctd<-subset(yflarv.ctd,doy>220)

str_name<-"./Environmental Data/expanded_BS_bathy.tif"
bathy<-raster(str_name)

bathybath<-as.bathy(bathy)
```

These data have been trimmed. The larval data are constrained to depths less than 176 meters and day of year greater than 220. Larvae are linked to CTD-derived, $in$ $situ$ temperature and salinity measurements.

## Descriptive Information: 

```{r,echo=FALSE}
lat_range<-c(paste(round(min(yflarv.ctd$lat),digits=1),round(max(yflarv.ctd$lat),digits=1),sep="-"))
lon_range<-c(paste(round(min(yflarv.ctd$lon),digits=1),round(max(yflarv.ctd$lon),digits=1),sep=" to "))
doy_range<-c(paste(min(yflarv.ctd$doy),max(yflarv.ctd$doy),sep="-"))
depth_range<-c(paste(round(min(yflarv.ctd$bottom_depth),digits=1),
                     round(max(yflarv.ctd$bottom_depth),digits=1),sep="-"))

df<-data.frame(lat_range,lon_range,doy_range,depth_range)

kable(df,col.names=c("Lat Range","Lon Range","Day of Year Range",
                     "Bottom Depth Range"),
      caption="Descriptive Metrics for Yellowfin Sole Larval Data",align='c')
```

The following two plots show *the day of year distribution for positive yellowfin sole larval catch* (left) and *the year distribution for positive yellowfin sole larval catch* (right). 

```{r, fig.width=7.5,fig.height=3.5,fig.fullwidth=TRUE,fig.cap="Yellowfin Sole Larvae", echo=FALSE}
par(mfrow=c(1,2))
hist(yflarv.ctd$doy[yflarv.ctd$Cper10m2>0],xlab="Day of Year",ylab="Positive Catch Frequencies",main="")
hist(yflarv.ctd$year[yflarv.ctd$Cper10m2>0],xlab="Year",ylab="",main="")
```

\newpage

The following plots show yellowfin sole larval catch distributions (Catch per unit effort, or per 10m$^2$) across five year increments from 1997 to 2016 
```{r,echo=FALSE}
yfylarv<-yflarv.ctd%>%mutate(yearbin=case_when(year<=2002~'1997-2002',
                                               year<=2007&year>2002~'2003-2007',
                                               year<=2012&year>2007~'2008-2012',
                                               year<=2017&year>2012~'2013-2016'))

yfylpos<-yfylarv[yfylarv$Cper10m2>0,]
yfylpos$logcpue<-log(yfylpos$Cper10m2+1)
yfylz<-yfylarv[yfylarv$Cper10m2==0,]
```


```{r,fig.height=10,fig.width=8,fig.align='center',fig.fullwidth=T,echo=FALSE,warning=FALSE}
autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=yfylz,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=yfylpos,aes(x=lon,y=lat,size=logcpue),
                               fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous("Log (CPUE+1)",range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  scale_shape_manual(values=('No Catch'=8))+
  facet_wrap(~yearbin)+
  labs(title="Yellowfin Sole Larvae")+ylab("Latitude")+xlab("Longitude")
```

## Larval Generalized Additive Models: 

Now we'll move into the GAMs. The following code *is only necessary if the data were re-trimmed and new GAMs need to be run*. In this case, modify markdown document such that "{eval = TRUE}". The other model figures are marked as "eval = FALSE" if they, as of the last model run, do not produce the best model results. **Make sure to save the new models as RDS objects.** 

Yellowfin sole larvae were best explained by the bivariate salinity-temperature model, in which the spatial and temporal distribution of larvae were modelled in association with a smooth containing *in situ* salinity-temperature data. 

We begin with the base larval model: 
```{r,eval=T,echo=FALSE}
lv.base<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
               s(bottom_depth,k=5),
             data=yflarv.ctd,family=tw(link='log'),method='REML')

saveRDS(lv.base,"./GAM Models/yf_larvae_base.rds")
```

```{r}
lv.base<-readRDS("./GAM Models/yf_larvae_base.rds")
summary(lv.base)
AIC(lv.base)
```

```{r,eval=T,echo=FALSE}
lv.add.sal<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                  s(bottom_depth,k=5)+
                  s(salinity),data=yflarv.ctd,family=tw(link='log'),
                method='REML')

saveRDS(lv.add.sal,file="./GAM Models/yf_larvae_addsal.rds")
```

```{r,echo=T,include=FALSE}
lv.add.sal<-readRDS("./GAM Models/yf_larvae_addsal.rds")
summary(lv.add.sal)
```

```{r,eval=T,echo=FALSE}
lv.add.temp<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                   s(bottom_depth,k=5)+
                   s(temperature),data=yflarv.ctd,family=tw(link='log'),
                 method='REML')

saveRDS(lv.add.temp,file="./GAM Models/yf_larvae_addtemp.rds")
```

```{r,include=FALSE,echo=FALSE}
lv.add.temp<-readRDS("./GAM Models/yf_larvae_addtemp.rds")
summary(lv.add.temp)
```

Then additive temperature and salinity, in individual additive terms. This is the second-best performing model. 
```{r,eval=T,echo=FALSE}
lv.temp.sal<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                   s(bottom_depth,k=5)+
                   s(temperature)+s(salinity),data=yflarv.ctd,
                 family=tw(link='log'),method='REML')

saveRDS(lv.temp.sal,file="./GAM Models/yf_larvae_addtempsal.rds")
```

```{r}
lv.temp.sal<-readRDS("./GAM Models/yf_larvae_addtempsal.rds")
summary(lv.temp.sal)
AIC(lv.temp.sal)
```

And finally, the best performing model: the bivariate salinity-temperature additive term: 
```{r,eval=T,echo=FALSE}
lv.2d<-gam((Cper10m2+1)~factor(year)+s(lon,lat)+s(doy,k=7)+s(bottom_depth)+
             te(salinity,temperature),data=yflarv.ctd,family=tw(link='log'),
           method='REML')

saveRDS(lv.2d,file="./GAM Models/yf_larvae_2d.rds")
```

```{r}
lv.2d<-readRDS("./GAM Models/yf_larvae_2d.rds")
summary(lv.2d)
AIC(lv.2d)
```

The following plot is the predicted yellowfin sole larval biogeography based on the best performing model, the bivariate salinity-temperature GAM. Observations (log transformed, n+1) are shown as well. 
```{r,echo=FALSE}
nlat=120
nlon=120
latd=seq(min(yflarv.ctd$lat,na.rm=TRUE),max(yflarv.ctd$lat,na.rm=TRUE),length.out=nlat)
lond=seq(min(yflarv.ctd$lon,na.rm=TRUE),max(yflarv.ctd$lon,na.rm=TRUE),length.out=nlon)

grid.extent<-expand.grid(lond,latd)
names(grid.extent)<-c('lon','lat')

grid.extent$dist<-NA
for(k in 1:nrow(grid.extent)){
  dist<-distance.function(grid.extent$lat[k],grid.extent$lon[k],
                          yflarv.ctd$lat,yflarv.ctd$lon)
  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-as.numeric(2008)
grid.extent$doy<-as.numeric(median(yflarv.ctd$doy,na.rm=TRUE))
grid.extent$bottom_depth<-NA
grid.extent$bottom_depth<-as.numeric(median(yflarv.ctd$bottom_depth,na.rm=TRUE))
grid.extent$temperature<-as.numeric(mean(yflarv.ctd$temperature))
grid.extent$salinity<-as.numeric(mean(yflarv.ctd$salinity))
grid.extent$pred<-predict(lv.2d,newdata=grid.extent)
grid.extent$pred[grid.extent$dist>30000]<-NA 

symcol<-adjustcolor("grey28",alpha=0.5)
```

```{r,fig.align='center',fig.fullwidth=TRUE,fig.wdith=6.5,fig.height=10,echo=FALSE}
par(mai=c(1,1,0.5,0.9),mfrow=c(2,1))
image.plot(lond,latd,t(matrix(grid.extent$pred,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=range(yflarv.ctd$lon,na.rm=TRUE),ylim=range(yflarv.ctd$lat,na.rm=TRUE),
           main='Predicted Larval Biogeography, 2D Model',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("(log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)


image.plot(lond,latd,t(matrix(grid.extent$pred,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=range(yflarv.ctd$lon,na.rm=TRUE),ylim=range(yflarv.ctd$lat,na.rm=TRUE),
           main='Predicted Larval Biogeography, 2D Model',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("(log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
symbols(yflarv.ctd$lon[yflarv.ctd$Cper10m2>0],
        yflarv.ctd$lat[yflarv.ctd$Cper10m2>0],
        circles=log(yflarv.ctd$Cper10m2+1)[yflarv.ctd$Cper10m2>0],
        inches=0.1,bg=symcol,fg='black',add=T)
points(yflarv.ctd$lon[yflarv.ctd$Cper10m2==0],
        yflarv.ctd$lat[yflarv.ctd$Cper10m2==0],pch="+",col="white")
map("worldHires",fill=T,col="gainsboro",add=T)
```

```{r,fig.align='center',fig.fullwidth=TRUE,fig.width=6.5,fig.height=5,echo=FALSE}
par(mai=c(1,1,0.6,0.9))
plot(lv.2d,select=2,shade=T,seWithMean=TRUE,shade.col="lightgrey",
     xlab="Day of Year",ylab="Day of Year Effect")
abline(h=0,col="mistyrose4",lty=2,lwd=1.3)
```

With this bivariate model, we can also calculate the predicted anomalous larval catch (more or less than expected) on a salinity-temperature plot. This figure shows that prediction, with observed larval catch (log(n=1)) overlaid. 
```{r,echo=FALSE}
ntemp<-100
nsal<-100
tempd<-seq(min(yflarv.ctd$temperature,na.rm=TRUE),max(yflarv.ctd$temperature,na.rm=TRUE),length.out=ntemp)
sald<-seq(min(yflarv.ctd$salinity,na.rm=T),max(yflarv.ctd$salinity,na.rm=T),length.out=nsal)

grid.extent<-expand.grid(sald,tempd)
names(grid.extent)<-c('salinity','temperature')

grid.extent$dist<-NA
for(k in 1:nrow(grid.extent)){
  dist<-euclidean.distance(grid.extent$salinity[k],grid.extent$temperature[k],
                               yflarv.ctd$salinity,yflarv.ctd$temperature)
  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-as.numeric(2005)
grid.extent$lon<-as.numeric(median(yflarv.ctd$lon))
grid.extent$lat<-as.numeric(median(yflarv.ctd$lat))
grid.extent$doy<-as.numeric(median(yflarv.ctd$doy,na.rm=TRUE))
grid.extent$bottom_depth<-NA
grid.extent$bottom_depth<-as.numeric(median(yflarv.ctd$bottom_depth,na.rm=TRUE))
grid.extent$pred<-predict(lv.2d,newdata=grid.extent)
grid.extent$pred[grid.extent$dist>mean(grid.extent$dist)]<-NA #threshold based on means
```

```{r,echo=FALSE,fig.align='center',fig.width=6.5,fig.height=10,fig.fullwidth=TRUE}
par(mai=c(1,1,0.5,0.9),mfrow=c(2,1))
image.plot(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
           col=hcl.colors(100,"Lajolla"),xlab='Salinity (psu)',
           ylab=expression(paste("Temperature ("^0, 'C)')),
           xlim=range(yflarv.ctd$salinity,na.rm=T),ylim=range(yflarv.ctd$temperature,na.rm=T),
           main='',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("Anomalies in (log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
        levels=c(2.2,2.8),col='white',add=T)

image.plot(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
           col=hcl.colors(100,"Lajolla"),xlab='Salinity (psu)',
           ylab=expression(paste("Temperature ("^0, 'C)')),
           xlim=range(yflarv.ctd$salinity,na.rm=T),ylim=range(yflarv.ctd$temperature,na.rm=T),
           main='Larval Biogeography By Temperature and Salinity',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("(log(C/(10m"^2,')+1)')),legend.shrink=0.3)
symbols(yflarv.ctd$salinity[yflarv.ctd$Cper10m2>0],
        yflarv.ctd$temperature[yflarv.ctd$Cper10m2>0],
        circles=log(yflarv.ctd$Cper10m2+1)[yflarv.ctd$Cper10m2>0],
        inches=0.1,bg=symcol,fg='black',add=T)
points(yflarv.ctd$salinity[yflarv.ctd$Cper10m2==0],
        yflarv.ctd$temperature[yflarv.ctd$Cper10m2==0],pch="+",col="white")
```

Now I'll calculate a specific range of temperature, salinity, and both temperature and salinity to evaluate breadth of environmental tolerances. 

```{r,echo=FALSE}
grid.extent<-grid.extent[!is.na(grid.extent$pred),] #remove NAs to get proper percentage of predictions

clip<-as.numeric(0.6*dim(grid.extent)[[1]]) #define nrows = to 60% of prediction grid 

grid.sort<-grid.extent%>%arrange(desc(pred))

grid.clip<-grid.sort%>%slice_head(n=clip) #clip top 70% of prediction grid 
head(grid.clip)

write.csv(grid.clip,'./Ichthyo Data/Yfs_larv_clip_60.csv')
```

```{r,echo=FALSE,fig.align='center',fig.width=8,fig.height=6,fig.fullwidth=TRUE}
temp<-ggplot(grid.extent,aes(x=temperature,y=pred))+geom_point()+geom_smooth(method="loess")+
  theme_bw()+theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(title="Predictions vs. Temperature",x='Temperature',y='Prediction')+
  scale_x_continuous(breaks=round(seq(min(grid.extent$temperature),max(grid.extent$temperature),by=0.75),1))

sal<-ggplot(grid.extent,aes(x=salinity,y=pred))+geom_point()+geom_smooth(method="loess")+
  theme_bw()+labs(title="Predictions vs. Salinity",x='Salinity',y='Prediction')+
  scale_x_continuous(breaks=round(seq(min(grid.extent$salinity),max(grid.extent$salinity),by=0.5),1))

par(mai=c(1,1,0.5,0.9))
ggarrange(temp,sal)
```

```{r,echo=FALSE,fig.align='center',fig.width=9,fig.height=6,fig.fullwidth=TRUE}
temp1<-ggplot(grid.extent,aes(x=temperature,y=pred))+geom_bin2d(bins=60)+
  theme_bw()+theme(axis.text.x=element_text(angle=45,hjust=1.2))+
  labs(title="Predictions vs. Temperature",x='Temperature',y='Prediction')+
  scale_fill_continuous(type='viridis')+
  scale_x_continuous(breaks=round(seq(min(grid.extent$temperature),max(grid.extent$temperature),by=0.75),1))

sal1<-ggplot(grid.extent,aes(x=salinity,y=pred))+geom_bin2d(bins=60)+
  theme_bw()+
  labs(title="Predictions vs. Salinity",x='Salinity',y='Prediction')+
  scale_fill_continuous(type='viridis')+
  scale_x_continuous(breaks=round(seq(min(grid.extent$salinity),max(grid.extent$salinity),by=0.5),1))

par(mai=c(1,1,0.5,0.9))
ggarrange(temp1,sal1)
```


To again share the improvements of the best performing models from the base models, we can look at the AIC division produces. 
```{r,echo=FALSE}
bestdiv<-c(AIC(lv.2d)/AIC(lv.base))
second<-c(AIC(lv.2d)/AIC(lv.temp.sal))
stage<-"Larvae"

df<-data.frame(stage,bestdiv,second)
kable(df,col.names=c("","Best Divided By Base","Best Divided By Second Best"),
      caption="Model Power through AIC Comparisons, Yellowfin Sole",align='c')
```

Reduction in MSE (%):  
```{r,echo=FALSE}
var.ratio.lv<-(summary(lv.base)$scale-summary(lv.2d)$scale)/summary(lv.base)$scale
perc<-var.ratio.lv*100
print(perc)
```

