---
title: "Streamlined Alaska Plaice Analyses & Figures"
author: "Laura Vary"
date: "1/10/2022"
header-includes: 
  - \usepackage{placeins}
output: 
  pdf_document: 
    toc: yes
---

```{r setup, echo=FALSE,message=FALSE,warning=FALSE,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE,message=FALSE,warning=FALSE,include=FALSE}
setwd("C:/Users/varyl/Desktop/GitHub/BeringSeaFishes_GAM_Analyses")
library(knitr)
library(raster)
library(marmap)
library(maps)
library(mgcv)
library(colorRamps) 
library(RColorBrewer)
library(fields)
library(dplyr)
library(ggplot2)
library(ggmap)
library(ggpubr)
library(scales)
library(mapdata)
source('./Organized Thesis Analyses (Code)/distance.function.R')
source("./Organized Thesis Analyses (Code)/vis.gam_COLORS.R")
source("./Organized Thesis Analyses (Code)/euclidean.distance.R")
```

## Purpose of This Document: 

The purpose of this document is to streamline files and associated analyses for the creation of generalized additive models that investigate spawning behavior and larval biogeography among fishes in the Bering Sea. This is mainly an automation document, with the goal of minimizing the back-and-forth between code files should data need to be modified or analyses re-ran.  

## Loading Data: 

Alaska plaice: both egg and larval data are included for this species. Plaice spawn during April and May, live over 30 years, and transform to juveniles at standard lengths $\geq$ to 10.7 mm. 

```{r}
apsub<-read.csv(file='./Ichthyo Data/Cleaned_Cut_ApEggs.csv',header=TRUE,
                check.names=TRUE)

aplarv.ctd<-read.csv(file='./Ichthyo Data/Cleaned_Cut_ApLarv_wCTD.csv',
                     header=TRUE,check.names=TRUE)
aplarv.ctd<-subset(aplarv.ctd,doy>80&doy<182)

reg.sst<-read.csv('./Environmental Data/Mar_SST_RegionalIndex_NCEP_BS.csv',
                  header=TRUE,check.names=TRUE)

#for(i in 1:nrow(apsub)){
 # apsub$reg.SST[i]<-reg.sst$SST[reg.sst$year==apsub$year[i]]}

str_name<-"./Environmental Data/expanded_BS_bathy.tif"
bathy<-raster(str_name)
```

These data have been trimmed. The egg data are constrained to depths <151 meters; temporally, the egg data are constrained to above the 99th day of year and below the 182nd day of the year (temporally centered on the spawning period of plaice). The egg data are also joined to regional temperature indices for each year (the reg.sst dataset). The larval data are constrained to depths <151 meters, between 80 and 165 day of year, and are linked to CTD-derived, $in$ $situ$ temperature and salinity measurements. 

The regional temperature index data are constrained to (-180, -151) degrees W and (50.5, 67.5) degrees N and reflect the average March temperature for each year across that region. March temperatures are chosen to estimate the conditions spawning plaice may have experienced, roughly two months before eggs appear in the water column.  

##Descriptive Information: Alaska Plaice

```{r,echo=FALSE}
lat_range<-c(paste(round(min(apsub$lat),digits=1),round(max(apsub$lat),digits=1),sep="-"))
lon_range<-c(paste(round(min(apsub$lon),digits=1),round(max(apsub$lon),digits=1),sep=" to "))
doy_range<-c(paste(min(apsub$doy),max(apsub$doy),sep="-"))
depth_range<-c(paste(round(min(apsub$bottom_depth),digits=1),
                     round(max(apsub$bottom_depth),digits=1),sep="-"))

df<-data.frame(lat_range,lon_range,doy_range,depth_range)

kable(df,col.names=c("Lat Range","Lon Range","Day of Year Range",
                     "Bottom Depth Range"),
      caption="Descriptive Metrics for Alaska Plaice Egg Data",align='c')
```
```{r, echo=FALSE}
lat_range<-c(paste(round(min(aplarv.ctd$lat),digits=1),round(max(aplarv.ctd$lat),digits=1),sep="-"))
lon_range<-c(paste(round(min(aplarv.ctd$lon),digits=1),round(max(aplarv.ctd$lon),digits=1),sep=" to "))
doy_range<-c(paste(min(aplarv.ctd$doy),max(aplarv.ctd$doy),sep="-"))
depth_range<-c(paste(round(min(aplarv.ctd$bottom_depth),digits=1),
                     round(max(aplarv.ctd$bottom_depth),digits=1),sep="-"))

df<-data.frame(lat_range,lon_range,doy_range,depth_range)

kable(df,col.names=c("Lat Range","Lon Range","Day of Year Range",
                     "Bottom Depth Range"),
      caption="Descriptive Metrics for Alaska Plaice Larval Data",align='c')
```

The following three plots show *the day of year distribution for positive plaice egg catch*, *the year distribution for positive plaice egg catch*. 

```{r, fig.width=7.5,fig.height=3.5,fig.fullwidth=TRUE, fig.cap='Alaska Plaice Eggs',echo=FALSE}
par(mfrow=c(1,2))
hist(apsub$doy[apsub$Cper10m2>0],xlab="Day of Year",ylab="Positive Catch Frequencies",main="")
hist(apsub$year[apsub$Cper10m2>0],xlab="Year",ylab="",main="")
```

The following three plots show *the day of year distribution for positive plaice larval catch*, *the year distribution for positive plaice larval catch*.  
```{r, fig.width=7.5,fig.height=3.5,fig.fullwidth=TRUE, fig.cap='Alaska Plaice Larvae',echo=FALSE}
par(mfrow=c(1,2))
hist(aplarv.ctd$doy[aplarv.ctd$Cper10m2>0],xlab="Day of Year",ylab="Positive Catch Frequencies",main="")
hist(aplarv.ctd$year[aplarv.ctd$Cper10m2>0],xlab="Year",ylab="",main="")
```

The following plots show Plaice egg and larval catch distributions (Catch per unit effort, or per 10m$^2$) across five year increments from 1979 to 2016. 
```{r,echo=FALSE,eval=T}
apysub<-apsub%>%mutate(yearbin=case_when(year<=1984~'1979-1984',
                                         year<=1989&year>1984~'1985-1989',
                                         year<=1994&year>1989~'1990-1994',
                                         year<=1999&year>1994~'1995-1999',
                                         year<=2004&year>1999~'2000-2004',
                                         year<=2009&year>2004~'2005-2009',
                                         year<=2014&year>2009~'2010-2014',
                                         year<=2019&year>2014~'2015-2016'))
apypos<-apysub[apysub$Cper10m2>0,]
apyz<-apysub[apysub$Cper10m2==0,]
apypos$logcpue<-log(apypos$Cper10m2+1)

bathybath<-as.bathy(bathy)
```

```{r,fig.height=10,fig.width=8,fig.align='center',fig.fullwidth=T,echo=FALSE,warning=FALSE,eval=T}
autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=apyz,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=apypos,aes(x=lon,y=lat,size=logcpue),
                        fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous("Log (CPUE + 1)",range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  facet_wrap(~yearbin)+
  labs(title="Alaska Plaice",shape="Log (CPUE + 1)")+ylab("Latitude")+xlab("Longitude")

```


```{r,echo=FALSE,eval=T}
apylarv<-aplarv.ctd%>%mutate(yearbin=case_when(year<=2002~'1997-2002',
                                               year<=2007&year>2002~'2003-2007',
                                               year<=2012&year>2007~'2008-2012',
                                               year<=2017&year>2012~'2013-2016'))

apylpos<-apylarv[apylarv$Cper10m2>0,]
apylpos$logcpue<-log(apylpos$Cper10m2+1)
apylz<-apylarv[apylarv$Cper10m2==0,]
```

```{r,fig.height=10,fig.width=8,fig.align='center',fig.fullwidth=T,echo=FALSE,warning=FALSE,eval=T}
autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=apylz,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=apylpos,aes(x=lon,y=lat,size=logcpue),
                               fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous("Log (CPUE+1)",range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  scale_shape_manual(values=('No Catch'=8))+
  facet_wrap(~yearbin)+
  labs(title="Plaice Larvae")+ylab("Latitude")+xlab("Longitude")

```


Now we'll move into the GAMs. The following code *is only necessary if the data were re-trimmed and new GAMs need to be run*. In this case, modify markdown document such that "{eval = TRUE}". The other model figures are marked as "eval = FALSE" if they, as of the last model run, do not produce the best model results. **Make sure to save the new models as RDS objects.** 

Alaska plaice eggs were best explained by the threshold geography model, in which geographic distribution of eggs varied differently below and above 2.12 degrees Celsius. 

## Generalized Additive Models: Alaska Plaice Eggs

The base model formulation: 
```{r, eval=TRUE}
eg.base<-gam((Cper10m2+1)~factor(year)+s(lon,lat)+s(doy)+s(bottom_depth,k=5),
             data=apsub,family=tw(link='log'),method='REML')

plot(eg.base,shade=FALSE,page=1,seWithMean=TRUE,scheme=2,scale=0)

summary(eg.base)
AIC(eg.base)

saveRDS(eg.base,file="./GAM Models/ap_egg_base.rds")
```
```{r,echo=FALSE,include=FALSE,eval=FALSE}
eg.base<-readRDS("./GAM Models/ap_egg_base.rds")
summary(eg.base)
AIC(eg.base)
```

The variable-coefficient geography formulation (in which geographic egg distributions vary differently in relation to regional SST indices). 
```{r, eval=T,echo=FALSE}
vc.geo<-gam((Cper10m2+1)~factor(year)+s(lon,lat)+s(doy)+s(bottom_depth,k=5)+
              s(lon,lat,by=reg.SST),data=apsub,family=tw(link='log'),
            method='REML')

par(mfrow=c(1,2))
plot(vc.geo,select=1,scheme=2,too.far=0.025,shade=FALSE,
     seWithMean=TRUE,xlab='Longitude',ylab='Latitude',
     main='V-C ap Egg Flex Geo, Avg. Variation')
map("world",fill=T,col="snow4",add=T)
plot(vc.geo,select=4,scheme=2,too.far=0.025,shade=FALSE,
     xlab='Longitude',ylab='Latitude',seWithMean=TRUE,
     main='V-C Flex Geo, Deviation from Avg. Variation')
map("world",fill=T,col="snow4",add=T)

summary(vc.geo)
AIC(vc.geo)

saveRDS(vc.geo,file="./GAM Models/ap_egg_vc_geo.rds")
```
```{r,echo=FALSE,include=FALSE,eval=FALSE}
vc.geo<-readRDS("./GAM Models/ap_egg_vc_geo.rds")
summary(vc.geo)
AIC(vc.geo)
```

The variable-coefficient phenology formulation (in which temporal (phenological) distribution of eggs vary in relation to regional SST indices). 
```{r, eval=T}
vc.pheno<-gam((Cper10m2+1)~factor(year)+s(lon,lat)+s(doy)+s(bottom_depth,k=5)+
                s(doy,by=reg.SST),data=apsub,family=tw(link='log'),
              method='REML')

par(oma=c(1,1,1,0.5),mar=c(3,3,3,1.5))
plot(vc.pheno,select=2,main='Alaska Plaice VC Phenology, Eggs',seWithMean=TRUE,
     ylim=c(-25,11))
abline(h=0,col='mistyrose4',lty=2,lwd=1.3)
par(oma=c(1,1,1,0.5),mar=c(3,3,3,1.5),new=TRUE)
plot(vc.pheno,select=4,seWithMean=TRUE,shade=TRUE,shade.col=col,ylim=c(-25,11))
legend('topright',legend=c('Flexible Phenology Smooth','Deviation from Avg.Phenology'),
       col=c(NA,col),lwd=c(2,2),cex=0.8)
mtext(c("Day of Year","Anomalies in log(CPUE+1)"),side=c(1,2),line=2.5)

summary(vc.pheno)
AIC(vc.pheno)

saveRDS(vc.pheno,file="./GAM Models/ap_egg_vc_pheno.rds")
```
```{r,echo=FALSE,include=FALSE,eval=F}
vc.pheno<-readRDS("./GAM Models/ap_egg_vc_pheno.rds")

```

The threshold phenology model formulation (in which the temporal (phenological) distribution of eggs vary differently above and below a threshold temperature: 
```{r,echo=FALSE}
temps<-sort(unique(reg.sst$SST))
bd<-4
temps.in<-temps[bd:(length(temps)-bd)]
```

```{r,echo=FALSE, eval=T}
aic.pheno<-NA*(temps.in)
thr.pheno<-as.list(1:(length(temps.in)))

for(i in 1:length(temps.in)){
  apsub$th<-factor(apsub$reg.SST<=temps.in[i])
  thr.pheno[[i]]<-gam((Cper10m2+1)~factor(year)+
                        s(lon,lat)+
                        s(bottom_depth,k=5)+
                        s(doy,by=th),
                      data=apsub,family=tw(link='log'),method='REML')
  aic.pheno[i]<-AIC(thr.pheno[[i]])
}

best.index.phe<-order(aic.pheno)[1]
thr.pheno<-thr.pheno[[best.index.phe]]
best.index.phe

summary(thr.pheno)
AIC(thr.pheno)

saveRDS(aic.pheno,file="./GAM Models/ap_egg_aic_pheno_list.rds")
saveRDS(thr.pheno,file="./GAM Models/ap_egg_thr_pheno.rds")
saveRDS(temps.in,file="./GAM Models/ap_egg_temps_in_pheno.rds")
saveRDS(best.index.phe,file="./GAM Models/ap_egg_best_index_phe.rds")
```
```{r,echo=FALSE,include=FALSE,eval=F}
temps.in<-readRDS("./GAM Models/ap_egg_temps_in_pheno.rds") #same for geo and pheno models 
aic.pheno<-readRDS("./GAM Models/ap_egg_aic_geo_list.rds")
thr.pheno<-readRDS("./GAM Models/ap_egg_thr_pheno.rds")
best.index.phe<-readRDS("./GAM Models/ap_egg_best_index_phe.rds")
summary(thr.pheno)
AIC(thr.pheno)
```

```{r, fig.width=6,fig.height=4.5,fig.fullwidth=TRUE,echo=FALSE,eval=T}
par(oma=c(1,1,1,0.5),mar=c(3,3,3,1.5))
plot(thr.pheno,select=4,main='Alaska Plaice Phenology, Eggs',seWithMean=TRUE,
     ylim=c(-4.2,2))
abline(h=0,col='mistyrose4',lty=2,lwd=1.3)
par(oma=c(1,1,1,0.5),mar=c(3,3,3,1.5),new=TRUE)
plot(thr.pheno,select=3,seWithMean=TRUE,shade=TRUE,shade.col="skyblue",ylim=c(-4.2,2))
legend('topright',legend=c('Below','Above'),col=c(NA,"skyblue"),lwd=c(2,2),cex=0.8)
mtext(c("Day of Year","Anomalies in log(CPUE+1)"),side=c(1,2),line=2.5)
```

```{r, fig.width=6,fig.height=4.5,fig.fullwidth=TRUE,echo=FALSE,eval=T}
plot(temps.in,aic.pheno,type='b',lwd=2,ylim=range(c(AIC(eg.base),aic.pheno)),
     main='Temperature Threshold Flexible Phenology',ylab='AIC Index',
     xlab='Temperature (degC)')
abline(h=AIC(eg.base),lty=2,lwd=2,col='sienna3')
abline(v=temps.in[best.index.phe],lty=2,lwd=2,col='steelblue3')
```

The threshold geography model formulation (in which the geographic distribution of eggs vary differently above and below a threshold temperature: 
*This is the best model to explain Alaska plaice egg variation across years, as of 1/10/2021*. 
```{r,echo=FALSE,eval=T}
aic.geo<-NA*(temps.in)
thr.geo<-as.list(1:(length(temps.in)))

for(i in 1:length(temps.in)){
  apsub$th<-factor(apsub$reg.SST<=temps.in[i])
  
  thr.geo[[i]]<-  gam((Cper10m2+1)~factor(year)+s(doy)+s(bottom_depth,k=5)+
                      s(lon,lat,by=th),data=apsub,
                    family=tw(link='log'),method='REML')
  
  aic.geo[i]<-AIC(thr.geo[[i]])
}

best.index.geo<-order(aic.geo)[1]
thr.geo<-thr.geo[[best.index.geo]]
best.index.geo[[1]]

summary(thr.geo)
AIC(thr.geo)

saveRDS(aic.geo,file="./GAM Models/ap_egg_aic_geo_list.rds")
saveRDS(thr.geo,file="./GAM Models/ap_egg_thr_geo.rds")
saveRDS(best.index.geo,file="./GAM Models/ap_egg_best_index_geo.rds")
```
```{r,eval=F}
thr.geo<-readRDS("./GAM Models/ap_egg_thr_geo.rds")
best.index.geo<-readRDS("./GAM Models/ap_egg_best_index_geo.rds")
aic.geo<-readRDS("./GAM Models/ap_egg_aic_geo_list.rds")
summary(thr.geo)
AIC(thr.geo)
print(temps.in[[best.index.geo]])
```

To confirm that this is indeed the best model, we can compare AIC values across all five tested models. 
```{r, echo=FALSE}
aic.base<-AIC(eg.base)
aic.thrph<-AIC(thr.pheno)
aic.thrge<-AIC(thr.geo)
aic.vcph<-AIC(vc.pheno)
aic.vcgeo<-AIC(vc.geo)

aic.apegg<-data.frame('model'=c('Base','Threshold Pheno','Threshold Geo',
                                'VC Pheno','VC Geo'),
                      'AIC_value'=c(aic.base,aic.thrph,aic.thrge,
                                    aic.vcph,aic.vcgeo))
```

```{r,fig.height=4.5,fig.width=6,fig.align='center',echo=FALSE}
plot(c(1:5),aic.apegg$AIC_value,main='AP',
     col=hcl.colors(5,"Zissou 1"),
     pch=19,cex=2,ylab='AIC Score',xlab='')
grid(nx=5,ny=14,col="lightgray")
text(c(1:5),aic.apegg$AIC_value,labels=round(aic.apegg$AIC_value),pos=c(4,3,3,2,2))
legend("bottomleft",legend=c('Base','Thr. Phenology','Thr. Geography',
                             'V-C Phenology','V-C Geography'),
       col=hcl.colors(5,"Zissou 1"),
       lwd=3,lty=1,cex=0.7)
```

```{r, fig.width=12,fig.height=8,fig.fullwidth=TRUE,fig.align='center', echo=FALSE, eval=T}
par(mfrow=c(1,2))
plot(thr.geo,select=4,scheme=2,too.far=0.025,
     main=paste('Below',round(temps.in[best.index.geo],digits=3),sep=" "),
     shade=FALSE,seWithMean=TRUE,xlab='Longitude',ylab='Latitude')
map("world",fill=T,col="gainsboro",add=T)
plot(thr.geo,select=3,scheme=2,too.far=0.025,
     main=paste('Above',round(temps.in[best.index.geo],digits=3),sep=" "),
     shade=FALSE,seWithMean=TRUE,xlab='Longitude',ylab='Latitude')
map("world",fill=T,col="gainsboro",add=T)
```

This is the below threshold (2.12 deg C) and above threshold predicted geographical distribution of Alaska plaice eggs (based on the threshold geography model). 
```{r, echo=FALSE}
nlat=120
nlon=120
latd=seq(min(apsub$lat),max(apsub$lat),length.out=nlat) #center grid over study region 
lond=seq(min(apsub$lon),max(apsub$lon),length.out=nlon)

grid.extent<-expand.grid(lond,latd)
names(grid.extent)<-c('lon','lat')

grid.extent$dist<-NA
for(k in 1:nrow(grid.extent)){
  dist<-distance.function(grid.extent$lat[k],grid.extent$lon[k],
                          apsub$lat,apsub$lon)
  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-as.numeric(2007)
grid.extent$doy<-as.numeric(median(apsub$doy,na.rm=TRUE))
grid.extent$bottom_depth<-NA
grid.extent$bottom_depth<-median(apsub$bottom_depth,na.rm=TRUE)
grid.extent$reg.SST<-NA
grid.extent$th<-"TRUE"
grid.extent$reg.SST<-mean(apsub$reg.SST,na.rm=TRUE) 
grid.extent$pred<-predict(thr.geo,newdata=grid.extent)
grid.extent$pred[grid.extent$dist>30000]<-NA 
grid.extent$th<-"FALSE"
grid.extent$pred2<-predict(thr.geo,newdata=grid.extent)
grid.extent$pred2[grid.extent$dist>30000]<-NA
```

```{r,fig.height=10,fig.width=6,fig.align='center',echo=FALSE}
par(mai=c(1,1,0.5,0.9),mfrow=c(1,2))
image.plot(lond,latd,t(matrix(grid.extent$pred,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=c(-180,-155),ylim=c(52,63),main='',zlim=c(-1.5,8.5),
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-1.8,
           legend.lab=expression(paste("log(C/(10m"^2,')+1)')),legend.shrink=0.4)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)

image.plot(lond,latd,t(matrix(grid.extent$pred2,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=c(-180,-155),ylim=c(52,63),main='',zlim=c(-1.5,8.5),
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-1.8,
           legend.lab=expression(paste("log(C/(10m"^2,')+1)')),legend.shrink=0.4)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)
```

```{r, fig.fullwidth=TRUE,fig.align='center',echo=FALSE,fig.width=6,fig.height=4}
par(oma=c(0.5,1,0.25,0.5))
plot(thr.geo,select=1,main="",seWithMean=TRUE,shade=TRUE,
     shade.col="lightgrey",xlab="Day of Year",ylab="Day of Year Effect",ylim=c(-2.5,1.5))
abline(h=0,col='mistyrose4',lty=2,lwd=1.3)
```

```{r, fig.fullwidth=TRUE,fig.align='center',echo=FALSE,fig.width=6,fig.height=4}
plot(temps.in,aic.geo,type='b',lwd=2,ylim=range(c(AIC(eg.base),aic.geo)),
     main='',xlab=expression(paste("Temperature ("^0,'C)')),ylab='AIC Score')
abline(h=AIC(eg.base),lty=2,lwd=2,col='sienna3')
abline(v=temps.in[best.index.geo],lty=2,lwd=2,col='steelblue3')
legend("bottomleft",cex=0.8,lty=c(2,2),lwd=c(1.2,1.2),col=c('sienna3',
                                                            'steelblue3'),
       legend=c("Base Model","Best Model"))
```

With the threshold geography model, we can predict places where there is a significant *difference* in predictions when comparing the above threshold spatial prediction to the below threshold spatial prediction. This plot is calculated by subtracting the spatial distribution *below* the threshold from the prediction *above* the threshold. 
```{r,echo=FALSE}
grid.extent<-expand.grid(lond,latd)
names(grid.extent)<-c('lon','lat')

grid.extent$dist<-NA
for(k in 1:nrow(grid.extent)){
  dist<-distance.function(grid.extent$lat[k],grid.extent$lon[k],
                          apsub$lat,apsub$lon)
  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-2007
grid.extent$doy<-median(apsub$doy)
grid.extent$reg.SST<-mean(apsub$reg.SST[apsub$reg.SST<temps.in[best.index.geo]]) #threshold temp chosen by AIC values
grid.extent$th<-"TRUE"
grid.extent$bottom_depth<-median(apsub$bottom_depth,na.rm=T)
grid.extent$pred<-predict(thr.geo,newdata=grid.extent)
grid.extent$se<-predict(thr.geo,newdata=grid.extent,se=T)[[2]]
grid.extent$pred.u<-grid.extent$pred+1.96*grid.extent$se #95% CI here
grid.extent$pred.l<-grid.extent$pred-1.96*grid.extent$se
grid.extent$pred[grid.extent$dist>30000]<-NA #remove predictions that are too far from positive data values
grid.extent$reg.SST<-mean(apsub$reg.SST[apsub$reg.SST>temps.in[best.index.geo]])
grid.extent$th<-"FALSE"
grid.extent$pred2<-predict(thr.geo,newdata=grid.extent)
grid.extent$se2<-predict(thr.geo,newdata=grid.extent,se=T)[[2]]
grid.extent$pred2.u<-grid.extent$pred2+1.96*grid.extent$se
grid.extent$pred2.l<-grid.extent$pred2-1.96*grid.extent$se
grid.extent$diff<-grid.extent$pred2-grid.extent$pred #calculate difference between two regimes

grid.extent$sig.pos<-c(grid.extent$pred2.l>grid.extent$pred.u) #isolate areas where there is a higher predicted CPUE at a higher temperature
grid.extent$sig.neg<-c(grid.extent$pred2.u<grid.extent$pred.l)
grid.extent$pos.diff<-grid.extent$diff*grid.extent$sig.pos #calculate areas with a significant positive difference at a higher temperature
grid.extent$neg.diff<-grid.extent$diff*grid.extent$sig.neg
max.slope<-max(grid.extent$diff,na.rm=T)
```

```{r,fig.align='center',fig.height=3.67,fig.width=5,echo=FALSE}
par(mai=c(1,1,0.5,0.5))
image.plot(lond,latd,t(matrix(grid.extent$diff,nrow=length(latd),ncol=length(lond),byrow=T)),
           col=hcl.colors(100,"BuPu",rev=T),ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')), 
           xlim=c(-180,-155),ylim=c(52,63),
           main="",
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("Change in log(C/(10m"^2,')+1)')),
           legend.shrink=0.4)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)
```

Reduction in MSE (%):  
```{r,echo=FALSE}
var.ratio.geo<-(summary(eg.base)$scale-summary(thr.geo)$scale)/summary(eg.base)$scale
perc<-var.ratio.geo*100
print(perc)
```


## Larval Generalized Additive Models: 
The following code *is only necessary if the data were re-trimmed and new GAMs need to be run*. In this case, modify markdown document such that "{eval = TRUE}". The other model figures are marked as "eval = FALSE" if they, as of the last model run, 
do not produce the best model results. These models are produced using conductivity-temperature-depth derived temperature and salinity measurements. 

We begin with the base larval model: 
```{r,eval=F}
lv.base<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
               s(bottom_depth,k=5),
             data=aplarv.ctd,family=tw(link='log'),method='REML')

saveRDS(lv.base,file="./GAM Models/ap_larval_base.rds")
```
```{r,include=FALSE}
lv.base<-readRDS("./GAM Models/ap_larval_base.rds")
summary(lv.base)
AIC(lv.base)
```

Then we add in additive salinity: 
```{r,eval=F}
lv.add.sal<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                  s(bottom_depth,k=5)+
                  s(salinity),data=aplarv.ctd,family=tw(link='log'),
                method='REML')

saveRDS(lv.add.sal,file="./GAM Models/ap_larval_addsal.rds")
```
```{r,include=FALSE}
lv.add.sal<-readRDS("./GAM Models/ap_larval_addsal.rds")
summary(lv.add.sal)
AIC(lv.add.sal)
```

Then additive temperature: 
```{r,eval=F}
lv.add.temp<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                   s(bottom_depth,k=5)+
                   s(temperature),data=aplarv.ctd,family=tw(link='log'),
                 method='REML')

saveRDS(lv.add.temp,file="./GAM Models/ap_larval_addtemp.rds")
```
```{r,include=FALSE}
lv.add.temp<-readRDS("./GAM Models/ap_larval_addtemp.rds")
summary(lv.add.temp)
AIC(lv.add.temp)
```

Then additive temperature and salinity, in individual additive terms: 
```{r,eval=F}
lv.temp.sal<-gam((Cper10m2+1)~factor(year)+s(doy,k=7)+s(lon,lat)+
                   s(bottom_depth,k=5)+
                   s(temperature)+s(salinity),data=aplarv.ctd,
                 family=tw(link='log'),method='REML')

saveRDS(lv.temp.sal,file="./GAM Models/ap_larval_addtempsal.rds")
```
```{r,include=FALSE}
lv.temp.sal<-readRDS("./GAM Models/ap_larval_addtempsal.rds")
summary(lv.temp.sal)
AIC(lv.temp.sal)
```

And finally, the best performing model: the bivariate salinity-temperature additive term: 
```{r,eval=F}
lv.2d<-gam((Cper10m2+1)~factor(year)+s(lon,lat)+s(doy,k=7)+s(bottom_depth,k=5)+
             te(salinity,temperature),data=aplarv.ctd,family=tw(link='log'),
           method='REML')

saveRDS(lv.2d,file="./GAM Models/ap_larval_2d.rds")
```
```{r,echo=FALSE}
lv.2d<-readRDS("./GAM Models/ap_larval_2d.rds")
summary(lv.2d)
AIC(lv.2d)
```

To confirm that this is indeed the best model, we can compare AIC values across all five tested models. 
```{r, echo=FALSE}
aic.base<-AIC(lv.base)
aic.sal<-AIC(lv.add.sal)
aic.temp<-AIC(lv.add.temp)
aic.saltemp<-AIC(lv.temp.sal)
aic.2d<-AIC(lv.2d)

aic.aplarv<-data.frame('model'=c('Base','Add Sal','Add Temp',
                                'Sal and Temp','2-d Sal-Temp'),
                      'AIC_value'=c(aic.base,aic.sal,aic.temp,
                                    aic.saltemp,aic.2d))
```



```{r,echo=FALSE}
windows(width=6,height=5);par(mai=c(1,1,0.5,0.9))
plot(c(1:5),aic.aplarv$AIC_value,main='AP',
     col=hcl.colors(5,"Zissou 1"),
     pch=19,cex=2,ylab='AIC Score',xlab='')
grid(nx=5,ny=14,col="lightgray")
text(c(1:5),aic.aplarv$AIC_value,labels=round(aic.aplarv$AIC_value),pos=c(4,1,3,2,2))
legend("bottomleft",legend=c('Base','Add. SSS','Add. SST',
                             'Add. SSS, SST','Bivariate SSS-SST'),
       col=hcl.colors(5,"Zissou 1"),
       lwd=3,lty=1,cex=0.7)
```

The following plot is the predicted Alaska Plaice larval biogeography based on the best performing model, the bivariate salinity-temperature GAM, next to the same plot with log(CPUE+1) observations shown. 
```{r,echo=FALSE}
nlat=120
nlon=120
latd=seq(min(aplarv.ctd$lat,na.rm=TRUE),max(aplarv.ctd$lat,na.rm=TRUE),length.out=nlat)
lond=seq(min(aplarv.ctd$lon,na.rm=TRUE),max(aplarv.ctd$lon,na.rm=TRUE),length.out=nlon)

grid.extent<-expand.grid(lond,latd)
names(grid.extent)<-c('lon','lat')

grid.extent$dist<-NA
for(k in 1:nrow(grid.extent)){
  dist<-distance.function(grid.extent$lat[k],grid.extent$lon[k],
                          aplarv.ctd$lat,aplarv.ctd$lon)
  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-as.numeric(2016)
grid.extent$doy<-as.numeric(median(aplarv.ctd$doy,na.rm=TRUE))
grid.extent$bottom_depth<-NA
grid.extent$bottom_depth<-as.numeric(median(aplarv.ctd$bottom_depth,na.rm=TRUE))
grid.extent$temperature<-as.numeric(mean(aplarv.ctd$temperature))
grid.extent$salinity<-as.numeric(mean(aplarv.ctd$salinity))
grid.extent$pred<-predict(lv.2d,newdata=grid.extent)
grid.extent$pred[grid.extent$dist>30000]<-NA 

col<-adjustcolor('grey28',alpha=0.5)
```

```{r,fig.align='center',fig.height=10,fig.width=6,echo=FALSE}
par(mai=c(1,1,0.5,0.9),mfrow=c(2,1))
image.plot(lond,latd,t(matrix(grid.extent$pred,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=c(-180,-155),ylim=c(52,63),
           main='',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)

image.plot(lond,latd,t(matrix(grid.extent$pred,nrow=length(latd),
                              ncol=length(lond),byrow=T)),col=hcl.colors(100,"Lajolla"),
           ylab=expression(paste("Latitude ("^0,'N)')),xlab=expression(paste("Longitude ("^0,'E)')),
           xlim=c(-180,-155),ylim=c(52,63),
           main='',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(bathy,levels=-c(50,200),labcex=0.4,col='grey28',add=T)
points(aplarv.ctd$lon[aplarv.ctd$Cper10m2==0],aplarv.ctd$lat[aplarv.ctd$Cper10m2==0],pch='+',col='white')
symbols(aplarv.ctd$lon[aplarv.ctd$Cper10m2>0],
        aplarv.ctd$lat[aplarv.ctd$Cper10m2>0],
        circles=log(aplarv.ctd$Cper10m2+1)[aplarv.ctd$Cper10m2>0],
        inches=0.1,bg=col,fg='black',add=T)
map("worldHires",fill=T,col="gainsboro",add=T)

```

```{r,fig.align='center',fig.fullwidth=TRUE,fig.width=5,fig.height=5,echo=FALSE}
par(mai=c(1,1,0.5,0.9))
plot(lv.2d,select=2,main="",xlab="Day of Year",
     ylab="Day of Year Effect",seWithMean=TRUE,shade=T,shade.col="lightgrey")
abline(h=0,lty=2,lwd=1.3,col="mistyrose4")
```

With this bivariate model, we can also calculate the predicted anomalous larval catch (more or less than expected) on a salinity-temperature plot. This figure shows that prediction, with observed larval catch (log(n=1)) overlaid. 
```{r,echo=FALSE}
ntemp<-100
nsal<-100
tempd<-seq(min(aplarv.ctd$temperature,na.rm=TRUE),max(aplarv.ctd$temperature,na.rm=TRUE),length.out=ntemp)
sald<-seq(min(aplarv.ctd$salinity,na.rm=T),max(aplarv.ctd$salinity,na.rm=T),length.out=nsal)

grid.extent<-expand.grid(sald,tempd)
names(grid.extent)<-c('salinity','temperature')

for(k in 1:nrow(grid.extent)){
  dist<-euclidean.distance(grid.extent$salinity[k],grid.extent$temperature[k],
                               aplarv.ctd$salinity,aplarv.ctd$temperature)

  grid.extent$dist[k]<-min(dist)
}

grid.extent$year<-as.numeric(2005)
grid.extent$lon<-as.numeric(median(aplarv.ctd$lon))
grid.extent$lat<-as.numeric(median(aplarv.ctd$lat))
grid.extent$doy<-as.numeric(median(aplarv.ctd$doy,na.rm=TRUE))
grid.extent$bottom_depth<-NA
grid.extent$bottom_depth<-as.numeric(median(aplarv.ctd$bottom_depth,na.rm=TRUE))
grid.extent$pred<-predict(lv.2d,newdata=grid.extent)
grid.extent$pred[grid.extent$dist>mean(grid.extent$dist)]<-NA #threshold is mean value from summary(grid.extent$dist)
```

```{r,echo=FALSE,fig.align='center',fig.width=6,fig.height=10,fig.fullwidth=TRUE}
par(mai=c(1,1,0.5,0.9),mfrow=c(2,1))
image.plot(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
           col=hcl.colors(100,"Lajolla"),xlab='Salinity (psu)',
           ylab=expression(paste("Temperature ("^0, 'C)')),
           xlim=c(29,33.5),ylim=c(-1.8,14),
           main='',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("log(C/(10m"^2,')+1)')),legend.shrink=0.3)
contour(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
        col="white",levels=c(1.4,1.9),add=T)

image.plot(sald,tempd,t(matrix(grid.extent$pred,nrow=length(tempd),ncol=length(sald),byrow=T)),
           col=hcl.colors(100,"Lajolla"),xlab='Salinity (psu)',
           ylab=expression(paste("Temperature ("^0, 'C)')),
           xlim=c(29,33.5),ylim=c(-1.8,14),
           main='Larval Biogeography By Temperature and Salinity',
           cex.main=1,cex.lab=1,cex.axis=0.9,legend.line=-2,
           legend.lab=expression(paste("log(C/(10m"^2,')+1)')),legend.shrink=0.3)
symbols(aplarv.ctd$salinity[aplarv.ctd$Cper10m2>0],
        aplarv.ctd$temperature[aplarv.ctd$Cper10m2>0],
        circles=log(aplarv.ctd$Cper10m2+1)[aplarv.ctd$Cper10m2>0],
        inches=0.1,bg=col,fg='black',add=T)
points(aplarv.ctd$salinity[aplarv.ctd$Cper10m2==0],aplarv.ctd$temperature[aplarv.ctd$Cper10m2==0],
       pch='+',col='white')
```

Now I'll calculate a specific range of temperature, salinity, and both temperature and salinity to evaluate breadth of environmental tolerances. 

For the univariate predictions, make a grid that holds either temperature or salinity constant depending on the variable of interest. The temperature and salinity values held constant were chosen based on the bivariate temperature,salinity plots; I attempted to capture a representative snapshot of variability in the univariate dimension. 

First, I'll determine the sum of larval log(CPUE+1) predictions and 60% of that sum, to know how many rows to cut from the larger extent. 

```{r,echo=FALSE}
sum<-sum(grid.extent$pred,na.rm=T)
sum

clip60<-0.6*sum
clip60 #determine 60% of summed CPUE predictions 

grid.extent1<-grid.extent
grid.extent1$pos<-seq.int(nrow(grid.extent1))
grid.extent1<-grid.extent1%>%arrange(pred)
grid.extent1$consum<-cumsum(grid.extent1$pred)
grid.extent1<-grid.extent1[grid.extent1$consum>clip60,]

grid.clip<-grid.extent1%>%arrange(grid.extent1$pos)

#grid.sort$con.sum<-cumsum(grid.sort$pred) #calculate cumulative sums for each row 

#grid.clip<-grid.sort[grid.sort$con.sum<=clip60,] #truncate top get 60% extent 
#ntemp.cl<-211
#nsal.cl<-211
#tempd.cl<-seq(min(aplarv.ctd$temperature,na.rm=TRUE),max(aplarv.ctd$temperature,na.rm=TRUE),length.out=ntemp.cl)
#sald.cl<-seq(min(aplarv.ctd$salinity,na.rm=T),max(aplarv.ctd$salinity,na.rm=T),length.out=nsal.cl)

percbi<-(dim(grid.clip[!is.na(grid.clip$pred),])[[1]]/dim(grid.extent)[[1]])*100
percbi 

#write.csv(grid.clip,'Ak_larv_clip_60.csv')
```

So this percentage (percbi, 38.8%) is the percentage of the total grid extent (in terms of nrow) within which 60% of the predicted log(CPUE+1) lies. I.e., 38.8% percent of the whole grid extent encompasses 60% of the predicted observations. 

```{r,echo=FALSE}
ntemp<-100
tempd<-seq(min(aplarv.ctd$temperature,na.rm=TRUE),max(aplarv.ctd$temperature,na.rm=TRUE),length.out=ntemp)

unit<-data.frame(matrix(ncol=1,nrow=100))
names(unit)<-'temperature'
unit$temperature<-as.numeric(tempd)

for(k in 1:nrow(unit)){
  unit$dist[k]<-unit$temperature[k]-aplarv.ctd$temperature[k]
}

unit$year<-as.numeric(2005)
unit$lon<-as.numeric(median(aplarv.ctd$lon))
unit$lat<-as.numeric(median(aplarv.ctd$lat))
unit$doy<-as.numeric(median(aplarv.ctd$doy,na.rm=TRUE))
unit$bottom_depth<-as.numeric(median(aplarv.ctd$bottom_depth,na.rm=TRUE))
unit$salinity<-31.25 #near the peak predictions 
unit$pred<-predict(lv.2d,newdata=unit)
unit$salinity<-31
unit$pred.1<-predict(lv.2d,newdata=unit)
unit$salinity<-30.75
unit$pred.2<-predict(lv.2d,newdata=unit)
unit$salinity<-30.25
unit$pred.3<-predict(lv.2d,newdata=unit)
unit$salinity<-31.5
unit$pred.4<-predict(lv.2d,newdata=unit)
unit$salinity<-31.75
unit$pred.5<-predict(lv.2d,newdata=unit)
unit$pred[unit$dist>mean(unit$dist)]<-NA
unit$pred.1[unit$dist>mean(unit$dist)]<-NA
unit$pred.2[unit$dist>mean(unit$dist)]<-NA
unit$pred.3[unit$dist>mean(unit$dist)]<-NA
unit$pred.4[unit$dist>mean(unit$dist)]<-NA
unit$pred.5[unit$dist>mean(unit$dist)]<-NA

nsal<-100
sald<-seq(min(aplarv.ctd$salinity,na.rm=T),max(aplarv.ctd$salinity,na.rm=T),length.out=nsal)
unis<-data.frame(matrix(ncol=1,nrow=100))
names(unis)<-'salinity'
unis$salinity<-as.numeric(sald)

for(k in 1:nrow(unis)){
  unis$dist[k]<-unis$salinity[k]-aplarv.ctd$salinity[k]
}

unis$year<-as.numeric(2005)
unis$lon<-as.numeric(median(aplarv.ctd$lon))
unis$lat<-as.numeric(median(aplarv.ctd$lat))
unis$doy<-as.numeric(median(aplarv.ctd$doy,na.rm=TRUE))
unis$bottom_depth<-as.numeric(median(aplarv.ctd$bottom_depth,na.rm=TRUE))
unis$temperature<-5 #centralized in peak predictions
unis$pred<-predict(lv.2d,newdata=unis)
unis$temperature<-5.25
unis$pred.1<-predict(lv.2d,newdata=unis)
unis$temperature<-5.75
unis$pred.2<-predict(lv.2d,newdata=unis)
unis$temperature<-4.75
unis$pred.3<-predict(lv.2d,newdata=unis)
unis$temperature<-4.25
unis$pred.4<-predict(lv.2d,newdata=unis)
unis$temperature<-6
unis$pred.5<-predict(lv.2d,newdata=unis)
unis$pred[unis$dist>mean(unis$dist)]<-NA
unis$pred.1[unis$dist>mean(unis$dist)]<-NA
unis$pred.2[unis$dist>mean(unis$dist)]<-NA
unis$pred.3[unis$dist>mean(unis$dist)]<-NA
unis$pred.4[unis$dist>mean(unis$dist)]<-NA
unis$pred.5[unis$dist>mean(unis$dist)]<-NA

unit<-unit[!is.na(unit$pred),] #the following clip will be based on the main prediction (pred), not the secondary ones (1-5)
unis<-unis[!is.na(unis$pred),]

clipt60<-0.6*(sum(unit$pred,na.rm=T))
clips60<-0.6*(sum(unis$pred,na.rm=T))

grid.t1<-unit
grid.t1$pos<-seq.int(nrow(grid.t1))
grid.t1<-grid.t1%>%arrange(pred)
grid.t1$consum<-cumsum(grid.t1$pred)
grid.t1<-grid.t1[grid.t1$consum>clipt60,]
grid.clip.t<-grid.t1%>%arrange(grid.t1$pos)

grid.s1<-unis
grid.s1$pos<-seq.int(nrow(grid.s1))
grid.s1<-grid.s1%>%arrange(pred)
grid.s1$consum<-cumsum(grid.s1$pred)
grid.s1<-grid.s1[grid.s1$consum>clips60,]
grid.clip.s<-grid.s1%>%arrange(grid.s1$pos)

#grid.sort.t<-unit%>%arrange(desc(pred))
#grid.sort.s<-unis%>%arrange(desc(pred))

#sumt<-sum(unit$pred,na.rm=T)
#sums<-sum(unis$pred,na.rm=T)

#clipt<-0.6*sumt
#clips<-0.6*sums

#grid.sort.t$con.sum<-cumsum(grid.sort.t$pred)
#grid.sort.s$con.sum<-cumsum(grid.sort.s$pred)

#grid.clip.t<-grid.sort.t[grid.sort.t$con.sum<=clipt,]
#grid.clip.s<-grid.sort.s[grid.sort.s$con.sum<=clips,]
```

```{r,echo=FALSE,fig.align='center',fig.width=8,fig.height=6,fig.fullwidth=TRUE,warning=FALSE}
temp<-ggplot(grid.clip.t,aes(x=temperature,y=pred))+geom_point(color='grey')+geom_smooth(method="loess")+
  geom_line(aes(x=temperature,y=pred.1),linetype="solid",color="grey48")+
  geom_line(aes(x=temperature,y=pred.2),linetype="twodash",color="grey48")+
  geom_line(aes(x=temperature,y=pred.3),linetype="longdash",color="grey48")+
  geom_line(aes(x=temperature,y=pred.4),linetype="dotted",color="grey48")+
  geom_line(aes(x=temperature,y=pred.5),linetype="dotdash",color="grey48")+
  theme_bw()+theme(axis.text.x=element_text(angle=45,hjust=1,size=7.5))+
  labs(x=expression(paste("Temperature ("^0,'C)')),y='Prediction')+
  scale_x_continuous(breaks=seq(-1.7,14,by=1),limits=c(-1.7,14))+
  scale_y_continuous(breaks=seq(1,7.3,by=1),limits=c(1,7.3))

sal<-ggplot(grid.clip.s,aes(x=salinity,y=pred))+geom_point(color='grey')+geom_smooth(method="loess")+
  geom_line(aes(x=salinity,y=pred.1),linetype="solid",color="grey48")+
  geom_line(aes(x=salinity,y=pred.2),linetype="twodash",color="grey48")+
  geom_line(aes(x=salinity,y=pred.3),linetype="longdash",color="grey48")+
  geom_line(aes(x=salinity,y=pred.4),linetype="dotted",color="grey48")+
  geom_line(aes(x=salinity,y=pred.5),linetype="dotdash",color="grey48")+
  theme_bw()+labs(x='Salinity (psu)',y='Prediction')+
  theme(axis.text.x=element_text(hjust=1,angle=45,size=7.5))+
  scale_x_continuous(breaks=round(seq(29,33.3,by=0.5),1),limits=c(29,33.3))+
  scale_y_continuous(breaks=seq(1,7.3,by=1),limits=c(1,7.3))

par(mai=c(1,1,0.5,0.9))
ggarrange(temp,sal)
```

For the bivariate analysis: 

```{r,echo=FALSE,fig.align='center',fig.width=6,fig.height=6,fig.fullwidth=TRUE}
par(mai=c(1,1,0.5,0.9))
bivar<-ggplot()+geom_tile(data=grid.clip,aes(x=salinity,y=temperature,fill=pred))+
  geom_vline(xintercept=31.25,linetype="dashed",color="black")+
  geom_vline(xintercept=31,linetype="solid",color="grey48")+
  geom_vline(xintercept=30.75,linetype="twodash",color="grey48")+
  geom_vline(xintercept=30.25,linetype="longdash",color="grey48")+
  geom_vline(xintercept=31.5,linetype="dotted",color="grey48")+
  geom_vline(xintercept=31.75,linetype="dotdash",color="grey48")+
  geom_hline(yintercept=5,linetype="dashed",color="black")+
  geom_hline(yintercept=5.25,linetype="solid",color="grey48")+
  geom_hline(yintercept=5.75,linetype="twodash",color="grey48")+
  geom_hline(yintercept=4.75,linetype="longdash",color="grey48")+
  geom_hline(yintercept=4.25,linetype="dotted",color="grey48")+
  geom_hline(yintercept=6,linetype="dotdash",color="grey48")+
  labs(x='Salinity (psu)',y=expression(paste("Temperature ("^0,'C)')))+scale_fill_viridis(name='Prediction')+
  scale_x_continuous(breaks=seq(29,33.3,by=0.5),limits=c(29,33.3))+
  scale_y_continuous(breaks=seq(-1.7,14,by=1),limits=c(-1.7,14))+
  theme(axis.text.x=element_text(hjust=1,angle=45,size=7.5))
bivar
```

Determine area in T-S units over which 60% of observations are contained.

```{r,echo=FALSE}
width<-max(grid.extent$salinity)-min(grid.extent$salinity)
height<-max(grid.extent$temperature)-min(grid.extent$temperature)
area.all<-width*height

width.cl<-max(grid.clip$salinity)-min(grid.clip$salinity)
height.cl<-max(grid.clip$temperature)-min(grid.clip$temperature)
area.clip<-width.cl*height.cl

perc.area<-(area.clip/area.all)*100
perc.area
```

41.78% of the salinity-temperature area contains 60% of the total predicted log(CPUE+1). 

To again share the improvements of the best performing models from the base models, we can look at the AIC division produces.
```{r,echo=FALSE,fig.align='center',fig.width=6,fig.height=10,fig.fullwidth=TRUE}
bestdiv<-c((AIC(thr.geo)/AIC(eg.base)),(AIC(lv.2d)/AIC(lv.base)))
second<-c((AIC(thr.geo)/AIC(vc.geo)),(AIC(lv.2d)/AIC(lv.temp.sal)))
stage<-c("Eggs","Larvae")

df<-data.frame(stage,bestdiv,second)
kable(df,col.names=c("","Best Divided By Base","Best Divided By Second Best"),
      caption="Model Power through AIC Comparisons, Alaska plaice",align='c')
```

Reduction in MSE (%):  
```{r,echo=FALSE}
var.ratio.lv<-(summary(lv.base)$scale-summary(lv.2d)$scale)/summary(lv.base)$scale
perc<-var.ratio.lv*100
print(perc)
```

## Appendices: 

The following plots investigate the phenological peaks in egg distribution to evaluate any spatial differences in egg distribution among these three peaks. 

```{r,echo=FALSE}
apd1<-apsub[apsub$Cper10m2>0&apsub$doy>122&apsub$doy<129,]
apd1$logcpue<-log(apd1$Cper10m2+1)
apdz1<-apsub[apsub$Cper10m2==0&apsub$doy>122&apsub$doy<129,]

apd2<-apsub[apsub$Cper10m2>0&apsub$doy>141&apsub$doy<150,]
apd2$logcpue<-log(apd2$Cper10m2+1)
apdz2<-apsub[apsub$Cper10m2==0&apsub$doy>141&apsub$doy<150,]

apd3<-apsub[apsub$Cper10m2>0&apsub$doy>162&apsub$doy<174,]
apd3$logcpue<-log(apd3$Cper10m2+1)
apdz3<-apsub[apsub$Cper10m2==0&apsub$doy>162&apsub$doy<174,]

```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
set1<-autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=apdz1,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=apd1,aes(x=lon,y=lat,size=logcpue),
                        fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous(expression(paste("log(C/(10m"^2,')+1)')),range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  labs(title="DOY 123-128",
       shape=expression(paste("log(C/(10m"^2,')+1)')))+
  ylab(expression(paste("Latitude ("^o,')')))+xlab(expression(paste("Longitude ("^o,'E)')))

set2<-autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=apdz2,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=apd2,aes(x=lon,y=lat,size=logcpue),
                        fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous(expression(paste("log(C/(10m"^2,')+1)')),range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  labs(title="DOY 142-149",
       shape=expression(paste("log(C/(10m"^2,')+1)')))+
  ylab(expression(paste("Latitude ("^o,')')))+xlab(expression(paste("Longitude ("^o,'E)')))

set3<-autoplot.bathy(bathybath,geom=c("contour","raster"),coast=TRUE,color="grey")+
  scale_fill_gradientn("Depth (m)",values=scales::rescale(c(-7600,0,1,2321)),
                       colors=c("steelblue4","#C7E0FF","grey50","grey80"))+
  geom_point(data=apdz3,aes(x=lon,y=lat),fill="violetred4",shape=8)+
  geom_point(data=apd3,aes(x=lon,y=lat,size=logcpue),
                        fill="#993300",shape=21,alpha=0.8)+
  scale_size_continuous(expression(paste("log(C/(10m"^2,')+1)')),range=c(0,9),breaks=pretty_breaks(9))+
  theme(panel.grid.major=element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        legend.position="bottom",legend.direction="horizontal",
        legend.text=element_text(size=8))+
  labs(title="DOY 163-173",
       shape=expression(paste("log(C/(10m"^2,')+1)')))+
  ylab(expression(paste("Latitude ("^o,')')))+xlab(expression(paste("Longitude ("^o,'E)')))
```


```{r,fig.fullwidth=T,fig.align='center',fig.width=8,fig.height=10,echo=F}
ggarrange(set1,set2,set3,common.legend=TRUE,legend="bottom")
```














