---
title: "Temp-Sal Diagram Code"
author: "Laura Vary"
date: "2/3/2022"
header-includes: 
  - \usepackage{placeins}
output: 
  pdf_document: 
    toc: yes
---

```{r setup, include=FALSE,warning=FALSE,echo=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/varyl/Desktop/GitHub/BeringSeaFishes_GAM_Analyses")
library(marelac)
library(plot3D)
library(DescTools)
```

## KDE Plots for Temperature and Salinity: 
This code will produce kernel density estimation plots for each larval species across both temperature and salinity to better estimate central tendencies of T-S associations. 
```{r,echo=FALSE}
aplarv.ctd<-read.csv(file='./Ichthyo Data/Cleaned_Cut_ApLarv_wCTD.csv',
                     header=TRUE,check.names=TRUE)
aplarv.ctd<-subset(aplarv.ctd,doy>80&doy<182)

fhlarv.ctd<-read.csv(file='./Ichthyo Data/Cleaned_Cut_FhLarv_wCTD.csv',
                     header=TRUE,check.names=TRUE)
fhlarv.ctd<-subset(fhlarv.ctd,doy>99)

pclarv.ctd<-read.csv('./Ichthyo Data/Cleaned_Cut_PcLarv_wCTD.csv')
pclarv.ctd<-subset(pclarv.ctd,lat<60)
pclarv.ctd<-subset(pclarv.ctd,doy>99&doy<160)

nrslarv.ctd<-read.csv('./Ichthyo Data/Cleaned_Cut_NrsLarv_wCTD.csv')
nrslarv.ctd<-subset(nrslarv.ctd,doy>100&doy<165)
nrslarv.ctd<-subset(nrslarv.ctd,lat<62)

pklarv.ctd<-read.csv(file='./Ichthyo Data/Cleaned_Cut_PkLarv_wCTD.csv',
                     header=TRUE,check.names=TRUE)
pklarv.ctd<-subset(pklarv.ctd,doy>90&doy<165)

yflarv.ctd<-read.csv(file='./Ichthyo Data/Cleaned_Cut_YfLarv_wCTD.csv',
                     header=TRUE,check.names=TRUE)
yflarv.ctd<-subset(yflarv.ctd,doy>220)
```

First, the temperature plots. 
```{r,echo=FALSE}
aptden<-density(aplarv.ctd$temperature,
               weights=(log(aplarv.ctd$Cper10m2+1)/sum(log(aplarv.ctd$Cper10m2+1))),na.rm=T)
fhtden<-density(fhlarv.ctd$temperature,
               weights=(log(fhlarv.ctd$Cper10m2+1)/sum(log(fhlarv.ctd$Cper10m2+1))),na.rm=T)
pctden<-density(pclarv.ctd$temperature,
               weights=(log(pclarv.ctd$Cper10m2+1)/sum(log(pclarv.ctd$Cper10m2+1))),na.rm=T)
nrstden<-density(nrslarv.ctd$temperature,
               weights=(log(nrslarv.ctd$Cper10m2+1)/sum(log(nrslarv.ctd$Cper10m2+1))),na.rm=T)
pktden<-density(pklarv.ctd$temperature,
               weights=(log(pklarv.ctd$Cper10m2+1)/sum(log(pklarv.ctd$Cper10m2+1))),na.rm=T)
yftden<-density(yflarv.ctd$temperature,
               weights=(log(yflarv.ctd$Cper10m2+1)/sum(log(yflarv.ctd$Cper10m2+1))),na.rm=T)
```

```{r,echo=FALSE,fig.width=7,fig.height=8,fig.align='center'}
plot(aptden,type="o",lty=1,lwd=0.75,pch=19,col="steelblue4",xlab='Temperature (*C)',ylab='Density Estimation',
     main="Larvae Temperature KDE",xlim=c(-3,14),cex=0.8)
lines(fhtden,type="o",lty=2,lwd=0.75,pch=8,col="orange",cex=0.8)
lines(pctden,type="o",lty=3,lwd=0.75,pch=18,col="darkolivegreen3",cex=0.8)
lines(pktden,type="o",lty=4,lwd=0.75,pch=15,col="mediumpurple1",cex=0.8)
lines(nrstden,type="o",lty=5,lwd=0.75,pch=11,col="palevioletred3",cex=0.8)
lines(yftden,type="o",lty=6,lwd=0.75,pch=4,col="goldenrod3",cex=0.8)
legend(x="topleft",legend=c("Plaice","Flathead","Cod","Pollock","Northern Rock","Yellowfin"),
       col=c("steelblue4","orange","darkolivegreen3","mediumpurple1","palevioletred3",
             "goldenrod3"),pch=c(19,8,18,15,11,4),
       lty=1,lwd=2)
axis(side=1,at=(-2:15))
```

Then the salinity plots. 
```{r,echo=FALSE}
apsden<-density(aplarv.ctd$salinity,
               weights=(log(aplarv.ctd$Cper10m2+1)/sum(log(aplarv.ctd$Cper10m2+1))),na.rm=T)
fhsden<-density(fhlarv.ctd$salinity,
               weights=(log(fhlarv.ctd$Cper10m2+1)/sum(log(fhlarv.ctd$Cper10m2+1))),na.rm=T)
pcsden<-density(pclarv.ctd$salinity,
               weights=(log(pclarv.ctd$Cper10m2+1)/sum(log(pclarv.ctd$Cper10m2+1))),na.rm=T)
nrssden<-density(nrslarv.ctd$salinity,
               weights=(log(nrslarv.ctd$Cper10m2+1)/sum(log(nrslarv.ctd$Cper10m2+1))),na.rm=T)
pksden<-density(pklarv.ctd$salinity,
               weights=(log(pklarv.ctd$Cper10m2+1)/sum(log(pklarv.ctd$Cper10m2+1))),na.rm=T)
yfsden<-density(yflarv.ctd$salinity,
               weights=(log(yflarv.ctd$Cper10m2+1)/sum(log(yflarv.ctd$Cper10m2+1))),na.rm=T)
```

```{r,echo=FALSE,fig.width=7,fig.height=8,fig.align='center'}
plot(apsden,type="o",lty=1,lwd=0.75,pch=19,col="steelblue4",xlab='Salinity (psu)',ylab='Density Estimation',
     main="Larvae Salinity KDE",ylim=c(0,2),cex=0.8)
lines(fhsden,type="o",lty=2,lwd=0.75,pch=8,col="orange",cex=0.8)
lines(pcsden,type="o",lty=3,lwd=0.75,pch=18,col="darkolivegreen3",cex=0.8)
lines(pksden,type="o",lty=4,lwd=0.75,pch=15,col="mediumpurple1",cex=0.8)
lines(nrssden,type="o",lty=5,lwd=0.75,pch=11,col="palevioletred3",cex=0.8)
lines(yfsden,type="o",lty=6,lwd=0.75,pch=4,col="goldenrod3",cex=0.8)
legend(x="topleft",legend=c("Plaice","Flathead","Cod","Pollock","Northern Rock","Yellowfin"),
       col=c("steelblue4","orange","darkolivegreen3","mediumpurple1","palevioletred3",
             "goldenrod3"),pch=c(19,8,18,15,11,4),
       lty=1,lwd=2)
```

## T-S Diagram with all species: 
This needs some work, but will ultimately show the central tendencies of larvae to associate with certain temperature-salinity values. 
```{r,echo=FALSE,include=FALSE}
dat<-read.csv('./Environmental Data/All_CTD_Data_8302021.csv',header=TRUE,check.names=TRUE)
dat<-subset(dat,Depth<11,na.rm=T)
dat<-dat[dat$Temperature<14,]
dat<-dat[dat$Salinity>29&dat$Salinity<36,]


mint=min(dat$Temperature,na.rm=T)
maxt=max(dat$Temperature,na.rm=T)
mins=min(dat$Salinity,na.rm=T)
maxs=max(dat$Salinity,na.rm=T)

temp<-seq(from=mint,to=maxt,length.out=180)
sal<-seq(from=mins,to=maxs,length.out=180)

sigma.c<-outer(sal,temp,FUN=function(S,t)sw_dens(S=S,t=t)-1000)
sigma.c

colr<-adjustcolor("grey",alpha=0.3)
```

```{r,fig.height=10,fig.width=8,fig.align='center',fig.fullwidth=T,echo=FALSE}
par(mai=c(1,1,0.5,0.9))
contour2D(x=sal,y=temp,z=sigma.c,lwd=1.6,lty=3,main='Temperature-Salinity Diagram',col='black',
          xlab=expression('Salinity (psu)'),ylab=expression('Temperature ('*~degree*C*')'))
DrawEllipse(x=31.675,y=5.95,radius.x=0.15,radius.y=0.3,rot=0, border='steelblue4',col=colr,lwd=3)#x axis = salinity, y axis = temperature
DrawEllipse(x=31.925,y=5.85,radius.x=0.25,radius.y=1.7,rot=0,border='orange',col=colr,lwd=3)
DrawEllipse(x=31.675,y=5.05,radius.x=0.05,radius.y=0.5,rot=0,border='darkolivegreen3',col=colr,lwd=3)
DrawEllipse(x=32,y=5.6,radius.x=0.6,radius.y=1.2,border='mediumpurple1',col=colr,lwd=3)
DrawEllipse(x=31.925,y=5.25,radius.x=0.35,radius.y=1.5,border='palevioletred3',col=colr,lwd=3)
DrawEllipse(x=31.45,y=9.6,radius.x=0.5,radius.y=0.8,border='goldenrod3',col=colr,lwd=3)
legend(x="bottomleft",legend=c("Plaice","Flathead","Cod","Pollock","Northern Rock","Yellowfin"),
       col=c("steelblue4","orange","darkolivegreen3","mediumpurple1","palevioletred3",
             "goldenrod3"),pch=16,bg='snow1')
```






